---
title: "ThermalPhysiologyFigures"
format: html
editor: visual
---

## Figures and Exploratory Analysis

This code is intended to set the stage for further analyses.

1.  Load in data and clean it up

2.  Make preliminary plots:

    -   fish length distribution

    -   Plots for visualizing CTmax in 2024 vs 2025 for just treatment 2

    -   Stream temperature

3.  Basic data analysis:

    -   Check data for normality (probably going to need to do a Welch's ANOVA with Games-Howell posthoc test for pairwise comparisons)

    -   Is size distribution within locations different across years?

    -   Is size distribution across locations different within years?

    -   Is CTmax different within locations across years?

    -   Is CTmax different across locations within years?

## Part 1: Load in Data and Clean

### Libraries

```{r}
lapply(c("tidyverse", "lme4", "ggthemes","car","emmeans","brglm2", "ggthemes", 
         "rethinking", "usethis", "here", "mgcv", "gratia", "extrafont","rstatix"), 
       require, character.only=T)
```

### Data

#### Physiology

```{r}
comp <- read.csv(here("data/IRA.thermaltolerance.compileddata.csv"))
head(comp)
```

##### 1. Rename columns and check structure

I am going to rename the Location column to Watershed, since there are two Big Creek's. This will make it more straightforward later on with color palettes and clarity in figures.

```{r}
comp_clean = comp %>% 
  rename(Watershed = Location)
unique(comp_clean$Watershed)

# rename locations to match the watershed name
comp_clean = comp_clean %>% 
    mutate(Watershed = case_when(
    Watershed == "Big Sur" ~ "Big Creek", # Big Creek in Big Sur
    Watershed == "Napa" ~ "Napa River", # Napa River
    Watershed == "E Austin" ~ "Russian River", # Russian River
    TRUE~ Watershed # keep Scott Creek and Carmel River the same
  ))
```

```{r}
str(comp_clean)
comp_clean$Date = as.Date(comp$Date, format = "%m/%d/%Y") #convert date to date structure
# Treatment should be a factor
comp_clean$Treatment = as.factor(comp_clean$Treatment)

# check for NA values
sum(is.na(comp_clean))
# there is one NA value in the length column

# add a year column
comp_clean <- comp_clean %>% 
  mutate(Year = year(Date))
comp_clean$Year = as.factor(comp_clean$Year)
comp_clean$Mass_g <- as.numeric(comp_clean$Mass_g)

# Add a LocationYear column
comp_clean <- comp_clean %>% 
  unite(WatershedYear, Watershed, Year, sep = " ", remove= FALSE)

# add a sample size column
#comp %>% 
#  group_by(Location) %>% 
#  summarise(sample_size = n())

# calculate relative weight
comp_clean <- comp_clean %>%
  mutate(
    #Where are these weight-length parameters from? FishBase?
    log10Ws = -4.790 + 2.928 * log10(Length_mm), 
    #Why taking log here
    Ws = 10^log10Ws
  )

comp_clean <- comp_clean %>%
  mutate(
    Wr = (Mass_g / Ws)*100
  )

```

#### Treatment 2 data ONLY

This is the dataframe we are tentatively planning on using for just chapter 1. This will be more or less a purely physiological paper.

From Big Creek, Scott Creek, and Carmel River

```{r}
treat2 = comp_clean %>% 
  filter(Treatment == 2) %>% 
  filter(Watershed %in% c("Big Creek", "Carmel River", "Scott Creek"))
```

#### Stream Temperature

### Color Palettes

```{r}
# palette generated by Coolors
fig1_palette= c(
  "Big Creek"="#744253",
  "Scott Creek"="#496A81",
  "Carmel River"="#9BC53D")
```

## Part 2: Preliminary Plots

Because this paper will only use Big Creek, Scott Creek, and Carmel River, these plots will only focus on those three sites.

### Size distribution

```{r}
size_dist <- ggplot(treat2, aes(x = Length_mm, fill = Watershed))+
  geom_histogram(binwidth = 5, , color = "black")+
  scale_fill_manual(values = fig1_palette)+
   scale_y_continuous(
    expand = expansion(mult = c(0, 0))
  )+
  labs(x = "Length (mm)",
       y = "Count")+
  theme_bw()+
  facet_grid(Watershed~Year);size_dist

ggsave("output/facet.fishlength.png", plot= size_dist, height = 4, width = 4.5, units = "in")
```

#### Within locations, across years

```{r}
size_distbyyear <- ggplot(treat2, aes(x = Length_mm, fill = Watershed))+
  geom_histogram(binwidth = 5, , color = "black")+
  scale_fill_manual(values = fig1_palette)+
   scale_y_continuous(
    expand = expansion(mult = c(0, 0))
  )+
  labs(x = "Length (mm)",
       y = "Count")+
  theme_bw()+
  facet_wrap(~Year); size_distbyyear
```

```{r}
size_distbylocation <- ggplot(treat2, aes(x = Length_mm, fill = Watershed))+
  geom_histogram(binwidth = 5, , color = "black")+
  scale_fill_manual(values = fig1_palette)+
   scale_y_continuous(
    expand = expansion(mult = c(0, 0))
  )+
  labs(x = "Length (mm)",
       y = "Count")+
  theme_bw()+
  facet_wrap(~Watershed); size_distbylocation
```

### Critical Thermal Maximum

```{r}
treat2 %>% 
  group_by(WatershedYear) %>% 
  mutate(mean_C = mean(LOE.Temp_C)) %>% 
  ungroup() %>% 
  filter(!Treatment == 1) %>% 
  ggplot(aes(x = as.factor(Watershed), y = LOE.Temp_C, color = Watershed)) +

  geom_jitter(width = .25,
    size = 2.25,
    alpha = 0.5) +

  stat_summary(
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.8),
    size = 4.5,
    stroke = 0.6
  ) +

  facet_wrap(~Year) +

  scale_color_manual(values = fig1_palette) +

  labs(
    x = "Watershed",
    y = "LOE Temp (°C)"
  ) +

  theme_bw() +
  theme(
    text = element_text(family = "Times New Roman", size = 8),
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 7),
    strip.text = element_text(size = 7),
    panel.spacing = unit(0.4, "lines"),
    legend.position = "bottom",
    legend.spacing.y = unit(0.05, "cm"),
  legend.margin = margin(t = 0, b = 0, unit = "cm"),
  legend.box.margin = margin(0, 0, 0, 0)
  )+
  guides(
  color = guide_legend(nrow = 2, byrow = TRUE,
                       keywidth = 0.4,
    keyheight = 0.4)
)

# may need to make points smaller for when I save the figure for the manuscript
```

### Violin Plot

```{r}
violinplot <- treat2 %>% 
  group_by(WatershedYear) %>% 
  mutate(mean_C = mean(LOE.Temp_C)) %>% 
  ungroup() %>% 
  filter(!Treatment == 1) %>% 
  ggplot(aes(x = as.factor(Watershed), y = LOE.Temp_C, fill = Watershed)) +

  geom_violin(alpha = 0.5)+
  geom_jitter(width = .15,
    size = 1.75,
    alpha = 0.6) +

  facet_wrap(~Year) +

  
  scale_fill_manual(values = fig1_palette) +

  labs(
    x = "Watershed",
    y = "LOE Temp (°C)"
  ) +

  theme_bw() +
  theme(
    text = element_text(family = "Times New Roman", size = 8),
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 7),
    strip.text = element_text(size = 7),
    panel.spacing = unit(0.4, "lines"),
    legend.position = "bottom",
    legend.spacing.y = unit(0.05, "cm"),
  legend.margin = margin(t = 0, b = 0, unit = "cm"),
  legend.box.margin = margin(0, 0, 0, 0)
  )+
  guides(
  color = guide_legend(nrow = 2, byrow = TRUE,
                       keywidth = 0.4,
    keyheight = 0.4)
); violinplot


```

```{r}
ggsave("output/violinplot.png", plot = violinplot, height = 4, width = 4, units = "in")
```

## Basic Data Analysis of CTMax

Before proceeding with data analysis, I need to do a normality check.

### Welch's ANOVA and Games-Howell

#### Normality check for ANOVA

I need to check to see if the data violates the assumptions of normality required for an ANOVA.

Assumptions:

1.  each subject belongs to only group (no repeats)

2.  no significant outliers

3.  normality

4.  homogeneity of variances

Outliers:

```{r}
treat2 %>% 
  group_by(Watershed, Year) %>%
  identify_outliers(LOE.Temp_C)
```

Normality:

```{r}
model <- lm(LOE.Temp_C~Watershed, data = treat2)
par(mfrow=c(2,2)) #makea2x2matrixofplots
plot(model) #make4diagnosticplots
```

```{r}
shapiro.test(resid(model))
# passes Shapiro test!

summary(lm(resid(model)~poly(predict(model),2)))
# passes the non-linearity of residuals test

library(lmtest)
bptest(model)

# however, it is heteroskedastic
```

The above tested for normality of residuals.

```{r}
car::leveneTest(LOE.Temp_C ~ Watershed, data = treat2)
```

I use a basic Welch's ANOVA with a Games-Howell posthoc pairwise comparison. This will be done on two levels:

#### Across years

```{r}
oneway.test(LOE.Temp_C ~ Watershed, data = treat2, var.equal = FALSE)
```

```{r}
gh_acrossyears<- games_howell_test(treat2, LOE.Temp_C ~ Watershed)
summary(gh_acrossyears)
write.csv(gh_acrossyears, "output/acrossyears.gameshowelltest.csv",row.names = FALSE) 
```

##### Results Across Years:

What we see here is that Big Creek is statistically different from all sites (p\< 0.05), but Carmel River and Scott Creek are not significantly different (p = 0.39)

------------------------------------------------------------------------

#### Within Years

**Within 2024:**

```{r}
treat2 %>% 
  filter(Year == 2024) %>% 
  oneway.test(LOE.Temp_C ~ Watershed, data = ., var.equal = FALSE)

gh2024 <- treat2 %>% 
  filter(Year == 2024) %>%
  games_howell_test(., LOE.Temp_C ~ Watershed)
summary(gh2024)
write.csv(gh2024, "output/acrossyears.gameshowelltest.csv",row.names = FALSE) 
```

In 2024, Scott Creek had statistically higher thermal tolerance than Big Creek and Carmel River.

**Within 2025:**

```{r}
treat2 %>% 
  filter(Year == 2025) %>% 
  oneway.test(LOE.Temp_C ~ Watershed, data = ., var.equal = FALSE)

gh2025 <- treat2 %>% 
  filter(Year == 2025) %>%
  games_howell_test(., LOE.Temp_C ~ Watershed)
```

In 2025, Carmel River had statistically higher thermal tolerance than the other sites (p\<\<\<0.05). Scott Creek and Big Creek are not different from one another.

#### Results text from ANOVA:

We compared sites within years using a Welch's ANOVA and Games-Howell posthoc test (R package rstatix). In 2024, Scott Creek population was statistically significantly more thermally tolerant than the other two locations (p \< 0.05). In 2025, Carmel River fish had significantly higher thermal tolerance than the other locations (p\<0.05).

------------------------------------------------------------------------

#### Compare sites across years

```{r}
# Carmel River
treat2 %>% 
  filter(Watershed == "Carmel River") %>% 
  oneway.test(LOE.Temp_C ~ Year, data = ., var.equal = FALSE)

cr_gh <- treat2 %>% 
  filter(Watershed == "Carmel River") %>%
  games_howell_test(., LOE.Temp_C ~ Year);cr_gh

# Very significant!
```

```{r}
# Big Creek
treat2 %>% 
  filter(Watershed == "Big Creek") %>% 
  oneway.test(LOE.Temp_C ~ Year, data = ., var.equal = FALSE)

bc_gh <- treat2 %>% 
  filter(Watershed == "Big Creek") %>%
  games_howell_test(., LOE.Temp_C ~ Year);bc_gh

# not significant!
```

```{r}
# Scott Creek
treat2 %>% 
  filter(Watershed == "Scott Creek") %>% 
  oneway.test(LOE.Temp_C ~ Year, data = ., var.equal = FALSE)

sc_gh <- treat2 %>% 
  filter(Watershed == "Scott Creek") %>%
  games_howell_test(., LOE.Temp_C ~ Year);sc_gh
```

### Results of CTMax

The only location with between year differences is Carmel River. Why is that?

------------------------------------------------------------------------

## Plot of CTmax data with Significance

```{r}
library(ggpubr)
p <- treat2 %>% 
  filter(Year == 2024) %>% 
  group_by(Watershed) %>% 
  ungroup() %>% 
  ggplot(aes(x = as.factor(Watershed), y = LOE.Temp_C, fill = Watershed)) +

  geom_violin(alpha = 0.5)+
  geom_jitter(width = .15,
    size = 1.75,
    alpha = 0.6) +

  facet_wrap(~Year) +

  
  scale_fill_manual(values = fig1_palette) +

  labs(
    x = "Watershed",
    y = "LOE Temp (°C)"
  ) +

  theme_bw() +
  theme(
    text = element_text(family = "Times New Roman", size = 8),
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 7),
    strip.text = element_text(size = 7),
    panel.spacing = unit(0.4, "lines"),
    legend.position = "bottom",
    legend.spacing.y = unit(0.05, "cm"),
  legend.margin = margin(t = 0, b = 0, unit = "cm"),
  legend.box.margin = margin(0, 0, 0, 0)
  )+
  guides(
  color = guide_legend(nrow = 2, byrow = TRUE,
                       keywidth = 0.4,
    keyheight = 0.4)
);p

sig2024 <- gh2024 %>%
  filter(!p.adj.signif == "ns") %>% 
  mutate(y.position = c(31.75, 32))

p+
  stat_pvalue_manual(
    sig2024,
    label = "p.adj.signif",
    y.position = "y.position",
    inherit.aes = FALSE
  )
colnames(gh2024)
```

```{r}
# 2025


sc_gh <- treat2 %>% 
  filter(Watershed == "Scott Creek") %>%
  games_howell_test(., LOE.Temp_C ~ Year);sc_gh
p <- treat2 %>% 
  filter(Year == 2025) %>% 
  group_by(Watershed) %>% 
  ungroup() %>% 
  ggplot(aes(x = as.factor(Watershed), y = LOE.Temp_C, fill = Watershed)) +

  geom_violin(alpha = 0.5)+
  geom_jitter(width = .15,
    size = 1.75,
    alpha = 0.6) +

  facet_wrap(~Year) +

  
  scale_fill_manual(values = fig1_palette) +

  labs(
    x = "Watershed",
    y = "LOE Temp (°C)"
  ) +

  theme_bw() +
  theme(
    text = element_text(family = "Times New Roman", size = 8),
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 7),
    strip.text = element_text(size = 7),
    panel.spacing = unit(0.4, "lines"),
    legend.position = "bottom",
    legend.spacing.y = unit(0.05, "cm"),
  legend.margin = margin(t = 0, b = 0, unit = "cm"),
  legend.box.margin = margin(0, 0, 0, 0)
  )+
  guides(
  color = guide_legend(nrow = 2, byrow = TRUE,
                       keywidth = 0.4,
    keyheight = 0.4)
);p

sig2025 <- gh2025 %>%
  filter(!p.adj.signif == "ns") %>% 
  mutate(y.position = c(31.9, 32))

p+
  stat_pvalue_manual(
    sig2025,
    label = "p.adj.signif",
    y.position = "y.position",
    inherit.aes = FALSE
  )

```

```{r}
# Compare sites across year
treat2 %>% 
  oneway.test(LOE.Temp_C ~ WatershedYear, data = ., var.equal = FALSE)

wsy_gh <- treat2 %>% 
  games_howell_test(., LOE.Temp_C ~ WatershedYear);wsy_gh



p <- treat2 %>% 
  ggplot(aes(x = as.factor(WatershedYear), y = LOE.Temp_C, fill = Watershed, shape = Year)) +

  geom_violin(alpha = 0.5)+
  geom_jitter(width = .05,
    size = 1.25,
    alpha = 0.6) +
  
  scale_fill_manual(values = fig1_palette) +

  labs(
    x = "Watershed",
    y = "LOE Temp (°C)"
  ) +

  theme_bw() +
  theme(
    text = element_text(family = "Times New Roman", size = 8),
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 7),
    strip.text = element_text(size = 7),
    panel.spacing = unit(0.4, "lines"),
    legend.position = "bottom",
    legend.spacing.y = unit(0.05, "cm"),
  legend.margin = margin(t = 0, b = 0, unit = "cm"),
  legend.box.margin = margin(0, 0, 0, 0)
  )+
  guides(
  color = guide_legend(nrow = 2, byrow = TRUE,
                       keywidth = 0.4,
    keyheight = 0.4)
);p

sig_wsy_gh <- wsy_gh %>% 
  filter(!p.adj.signif == "ns") %>% 
  mutate(y.position = c(32, 31.9, 32.1, 32.2,32.3,32.4,32.5))

sigplot_ALL <- p+
  stat_pvalue_manual(
    sig_wsy_gh,
    label = "p.adj.signif",
    y.position = "y.position",
    inherit.aes = FALSE
  ); sigplot_ALL

ggsave("output/sigplot_ALL.png", plot = sigplot_ALL, height = 4, width = 5, units = "in")
```

## Basic Data Analysis of Length

### Normality check

To fit a linear model with length, I need to test the four assumptions of general linear models.

```{r}

```

## Stream Temperature

```{r}

```
