---
title: "ThermalPhysiologyFigures"
format: html
editor: visual
---

## Figures and Exploratory Analysis

This code is intended to set the stage for further analyses.

1.  Load in data and clean it up

2.  Make preliminary plots:

    -   fish length distribution

    -   Plots for visualizing CTmax in 2024 vs 2025 for just treatment 2

    -   Stream temperature

3.  Basic data analysis:

    -   Check data for normality (probably going to need to do a Welch's ANOVA with Games-Howell posthoc test for pairwise comparisons)

    -   Is size distribution within locations different across years?

    -   Is size distribution across locations different within years?

    -   Is CTmax different within locations across years?

    -   Is CTmax different across locations within years?

## Part 1: Load in Data and Clean

### Libraries

```{r}
lapply(c("tidyverse", "lme4", "ggthemes","car","emmeans","brglm2", "ggthemes", 
         "rethinking", "usethis", "here", "mgcv", "gratia", "extrafont","rstatix"), 
       require, character.only=T)
```

### Data

#### Physiology

```{r}
comp <- read.csv(here("data/IRA.thermaltolerance.compileddata.csv"))
head(comp)
```

##### 1. Rename columns and check structure

I am going to rename the Location column to Watershed, since there are two Big Creek's. This will make it more straightforward later on with color palettes and clarity in figures.

```{r}
comp_clean = comp %>% 
  rename(Watershed = Location)
unique(comp_clean$Watershed)

# rename locations to match the watershed name
comp_clean = comp_clean %>% 
    mutate(Watershed = case_when(
    Watershed == "Big Sur" ~ "Big Creek", # Big Creek in Big Sur
    Watershed == "Napa" ~ "Napa River", # Napa River
    Watershed == "E Austin" ~ "Russian River", # Russian River
    TRUE~ Watershed # keep Scott Creek and Carmel River the same
  ))
```

```{r}
str(comp_clean)
comp_clean$Date = as.Date(comp$Date, format = "%m/%d/%Y") #convert date to date structure
# Treatment should be a factor
comp_clean$Treatment = as.factor(comp_clean$Treatment)

# check for NA values
sum(is.na(comp_clean))
# there is one NA value in the length column

# add a year column
comp_clean <- comp_clean %>% 
  mutate(Year = year(Date))
comp_clean$Year = as.factor(comp_clean$Year)
comp_clean$Mass_g <- as.numeric(comp_clean$Mass_g)

# Add a LocationYear column
comp_clean <- comp_clean %>% 
  unite(WatershedYear, Watershed, Year, sep = " ", remove= FALSE)

# add a sample size column
#comp %>% 
#  group_by(Location) %>% 
#  summarise(sample_size = n())

# calculate relative weight
comp_clean <- comp_clean %>%
  mutate(
    #Where are these weight-length parameters from? FishBase?
    log10Ws = -4.790 + 2.928 * log10(Length_mm), 
    #Why taking log here
    Ws = 10^log10Ws
  )

comp_clean <- comp_clean %>%
  mutate(
    Wr = (Mass_g / Ws)*100
  )

```

#### Treatment 2 data ONLY

This is the dataframe we are tentatively planning on using for just chapter 1. This will be more or less a purely physiological paper.

From Big Creek, Scott Creek, and Carmel River

```{r}
treat2 = comp_clean %>% 
  filter(Treatment == 2) %>% 
  filter(Watershed %in% c("Big Creek", "Carmel River", "Scott Creek"))
```

### Color Palettes

```{r}
# palette generated by Coolors
fig1_palette= c(
  "Big Creek"="#744253",
  "Scott Creek"="#496A81",
  "Carmel River"="#9BC53D")
```

## Part 2: Preliminary Plots

Because this paper will only use Big Creek, Scott Creek, and Carmel River, these plots will only focus on those three sites.

### Size distribution

```{r}
size_dist <- ggplot(treat2, aes(x = Length_mm, fill = Watershed))+
  geom_histogram(binwidth = 5, , color = "black")+
  scale_fill_manual(values = fig1_palette)+
   scale_y_continuous(
    expand = expansion(mult = c(0, 0))
  )+
  labs(x = "Length (mm)",
       y = "Count")+
  theme_bw()+
  facet_grid(Watershed~Year);size_dist

ggsave("output/facet.fishlength.png", plot= size_dist, height = 4, width = 4.5, units = "in")
```

#### Within locations, across years

```{r}
size_distbyyear <- ggplot(treat2, aes(x = Length_mm, fill = Watershed))+
  geom_histogram(binwidth = 5, , color = "black")+
  scale_fill_manual(values = fig1_palette)+
   scale_y_continuous(
    expand = expansion(mult = c(0, 0))
  )+
  labs(x = "Length (mm)",
       y = "Count")+
  theme_bw()+
  facet_wrap(~Year); size_distbyyear
```

```{r}
size_distbylocation <- ggplot(treat2, aes(x = Length_mm, fill = Watershed))+
  geom_histogram(binwidth = 5, , color = "black")+
  scale_fill_manual(values = fig1_palette)+
   scale_y_continuous(
    expand = expansion(mult = c(0, 0))
  )+
  labs(x = "Length (mm)",
       y = "Count")+
  theme_bw()+
  facet_wrap(~Watershed); size_distbylocation
```

### Critical Thermal Maximum

```{r}
treat2 %>% 
  group_by(WatershedYear) %>% 
  mutate(mean_C = mean(LOE.Temp_C)) %>% 
  ungroup() %>% 
  filter(!Treatment == 1) %>% 
  ggplot(aes(x = as.factor(Watershed), y = LOE.Temp_C, color = Watershed)) +

  geom_jitter(width = .25,
    size = 2.25,
    alpha = 0.5) +

  stat_summary(
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.8),
    size = 4.5,
    stroke = 0.6
  ) +

  facet_wrap(~Year) +

  scale_color_manual(values = fig1_palette) +

  labs(
    x = "Watershed",
    y = "LOE Temp (°C)"
  ) +

  theme_bw() +
  theme(
    text = element_text(family = "Times New Roman", size = 8),
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 7),
    strip.text = element_text(size = 7),
    panel.spacing = unit(0.4, "lines"),
    legend.position = "bottom",
    legend.spacing.y = unit(0.05, "cm"),
  legend.margin = margin(t = 0, b = 0, unit = "cm"),
  legend.box.margin = margin(0, 0, 0, 0)
  )+
  guides(
  color = guide_legend(nrow = 2, byrow = TRUE,
                       keywidth = 0.4,
    keyheight = 0.4)
)

# may need to make points smaller for when I save the figure for the manuscript
```

### Violin Plot

```{r}
violinplot <- treat2 %>% 
  group_by(WatershedYear) %>% 
  mutate(mean_C = mean(LOE.Temp_C)) %>% 
  ungroup() %>% 
  filter(!Treatment == 1) %>% 
  ggplot(aes(x = as.factor(Watershed), y = LOE.Temp_C, fill = Watershed)) +

  geom_violin(alpha = 0.5)+
  geom_jitter(width = .15,
    size = 1.75,
    alpha = 0.6) +

  facet_wrap(~Year) +

  
  scale_fill_manual(values = fig1_palette) +

  labs(
    x = "Watershed",
    y = "LOE Temp (°C)"
  ) +

  theme_bw() +
  theme(
    text = element_text(family = "Times New Roman", size = 8),
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 7),
    strip.text = element_text(size = 7),
    panel.spacing = unit(0.4, "lines"),
    legend.position = "bottom",
    legend.spacing.y = unit(0.05, "cm"),
  legend.margin = margin(t = 0, b = 0, unit = "cm"),
  legend.box.margin = margin(0, 0, 0, 0)
  )+
  guides(
  color = guide_legend(nrow = 2, byrow = TRUE,
                       keywidth = 0.4,
    keyheight = 0.4)
); violinplot


```

```{r}
ggsave("output/violinplot.png", plot = violinplot, height = 4, width = 4, units = "in")
```

## Basic Data Analysis of CTMax

Before proceeding with data analysis, I need to do a normality check.

### Welch's ANOVA and Games-Howell

#### Normality check for ANOVA

I need to check to see if the data violates the assumptions of normality required for an ANOVA.

Assumptions:

1.  each subject belongs to only group (no repeats)

2.  no significant outliers

3.  normality

4.  homogeneity of variances

Outliers:

```{r}
treat2 %>% 
  group_by(Watershed, Year) %>%
  identify_outliers(LOE.Temp_C)
```

Normality:

```{r}
model <- lm(LOE.Temp_C~Watershed, data = treat2)
par(mfrow=c(2,2)) #makea2x2matrixofplots
plot(model) #make4diagnosticplots
```

```{r}
shapiro.test(resid(model))
# passes Shapiro test!

summary(lm(resid(model)~poly(predict(model),2)))
# passes the non-linearity of residuals test

library(lmtest)
bptest(model)

# however, it is heteroskedastic
```

The above tested for normality of residuals.

```{r}
car::leveneTest(LOE.Temp_C ~ Watershed, data = treat2)
```

I use a basic Welch's ANOVA with a Games-Howell posthoc pairwise comparison. This will be done on two levels:

#### Across years

```{r}
oneway.test(LOE.Temp_C ~ Watershed, data = treat2, var.equal = FALSE)
```

```{r}
gh_acrossyears<- games_howell_test(treat2, LOE.Temp_C ~ Watershed)
summary(gh_acrossyears)
write.csv(gh_acrossyears, "output/acrossyears.gameshowelltest.csv",row.names = FALSE) 
```

##### Results Across Years:

What we see here is that Big Creek is statistically different from all sites (p\< 0.05), but Carmel River and Scott Creek are not significantly different (p = 0.39)

------------------------------------------------------------------------

#### Within Years

**Within 2024:**

```{r}
treat2 %>% 
  filter(Year == 2024) %>% 
  oneway.test(LOE.Temp_C ~ Watershed, data = ., var.equal = FALSE)

gh2024 <- treat2 %>% 
  filter(Year == 2024) %>%
  games_howell_test(., LOE.Temp_C ~ Watershed)
summary(gh2024)
write.csv(gh2024, "output/acrossyears.gameshowelltest.csv",row.names = FALSE) 
```

In 2024, Scott Creek had statistically higher thermal tolerance than Big Creek and Carmel River.

**Within 2025:**

```{r}
treat2 %>% 
  filter(Year == 2025) %>% 
  oneway.test(LOE.Temp_C ~ Watershed, data = ., var.equal = FALSE)

gh2025 <- treat2 %>% 
  filter(Year == 2025) %>%
  games_howell_test(., LOE.Temp_C ~ Watershed)
```

In 2025, Carmel River had statistically higher thermal tolerance than the other sites (p\<\<\<0.05). Scott Creek and Big Creek are not different from one another.

#### Results text from ANOVA:

We compared sites within years using a Welch's ANOVA and Games-Howell posthoc test (R package rstatix). In 2024, Scott Creek population was statistically significantly more thermally tolerant than the other two locations (p \< 0.05). In 2025, Carmel River fish had significantly higher thermal tolerance than the other locations (p\<0.05).

------------------------------------------------------------------------

#### Compare sites across years

```{r}
# Carmel River
treat2 %>% 
  filter(Watershed == "Carmel River") %>% 
  oneway.test(LOE.Temp_C ~ Year, data = ., var.equal = FALSE)

cr_gh <- treat2 %>% 
  filter(Watershed == "Carmel River") %>%
  games_howell_test(., LOE.Temp_C ~ Year);cr_gh

# Very significant!
```

```{r}
# Big Creek
treat2 %>% 
  filter(Watershed == "Big Creek") %>% 
  oneway.test(LOE.Temp_C ~ Year, data = ., var.equal = FALSE)

bc_gh <- treat2 %>% 
  filter(Watershed == "Big Creek") %>%
  games_howell_test(., LOE.Temp_C ~ Year);bc_gh

# not significant!
```

```{r}
# Scott Creek
treat2 %>% 
  filter(Watershed == "Scott Creek") %>% 
  oneway.test(LOE.Temp_C ~ Year, data = ., var.equal = FALSE)

sc_gh <- treat2 %>% 
  filter(Watershed == "Scott Creek") %>%
  games_howell_test(., LOE.Temp_C ~ Year);sc_gh
```

### Results of CTMax

The only location with between year differences is Carmel River. Why is that?

------------------------------------------------------------------------

## Plot of CTmax data with Significance

```{r}
library(ggpubr)
p <- treat2 %>% 
  filter(Year == 2024) %>% 
  group_by(Watershed) %>% 
  ungroup() %>% 
  ggplot(aes(x = as.factor(Watershed), y = LOE.Temp_C, fill = Watershed)) +

  geom_violin(alpha = 0.5)+
  geom_jitter(width = .15,
    size = 1.75,
    alpha = 0.6) +

  facet_wrap(~Year) +

  
  scale_fill_manual(values = fig1_palette) +

  labs(
    x = "Watershed",
    y = "LOE Temp (°C)"
  ) +

  theme_bw() +
  theme(
    text = element_text(family = "Times New Roman", size = 8),
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 7),
    strip.text = element_text(size = 7),
    panel.spacing = unit(0.4, "lines"),
    legend.position = "bottom",
    legend.spacing.y = unit(0.05, "cm"),
  legend.margin = margin(t = 0, b = 0, unit = "cm"),
  legend.box.margin = margin(0, 0, 0, 0)
  )+
  guides(
  color = guide_legend(nrow = 2, byrow = TRUE,
                       keywidth = 0.4,
    keyheight = 0.4)
);p

sig2024 <- gh2024 %>%
  filter(!p.adj.signif == "ns") %>% 
  mutate(y.position = c(31.75, 32))

p+
  stat_pvalue_manual(
    sig2024,
    label = "p.adj.signif",
    y.position = "y.position",
    inherit.aes = FALSE
  )

```

```{r}
# 2025


sc_gh <- treat2 %>% 
  filter(Watershed == "Scott Creek") %>%
  games_howell_test(., LOE.Temp_C ~ Year);sc_gh
p <- treat2 %>% 
  filter(Year == 2025) %>% 
  group_by(Watershed) %>% 
  ungroup() %>% 
  ggplot(aes(x = as.factor(Watershed), y = LOE.Temp_C, fill = Watershed)) +

  geom_violin(alpha = 0.5)+
  geom_jitter(width = .15,
    size = 1.75,
    alpha = 0.6) +

  facet_wrap(~Year) +

  
  scale_fill_manual(values = fig1_palette) +

  labs(
    x = "Watershed",
    y = "LOE Temp (°C)"
  ) +

  theme_bw() +
  theme(
    text = element_text(family = "Times New Roman", size = 8),
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 7),
    strip.text = element_text(size = 7),
    panel.spacing = unit(0.4, "lines"),
    legend.position = "bottom",
    legend.spacing.y = unit(0.05, "cm"),
  legend.margin = margin(t = 0, b = 0, unit = "cm"),
  legend.box.margin = margin(0, 0, 0, 0)
  )+
  guides(
  color = guide_legend(nrow = 2, byrow = TRUE,
                       keywidth = 0.4,
    keyheight = 0.4)
);p

sig2025 <- gh2025 %>%
  filter(!p.adj.signif == "ns") %>% 
  mutate(y.position = c(31.9, 32))

p+
  stat_pvalue_manual(
    sig2025,
    label = "p.adj.signif",
    y.position = "y.position",
    inherit.aes = FALSE
  )

```

```{r}
# Compare sites across year
treat2 %>% 
  oneway.test(LOE.Temp_C ~ WatershedYear, data = ., var.equal = FALSE)

wsy_gh <- treat2 %>% 
  games_howell_test(., LOE.Temp_C ~ WatershedYear);wsy_gh



p <- treat2 %>% 
  ggplot(aes(x = as.factor(WatershedYear), y = LOE.Temp_C, fill = Watershed, shape = Year)) +

  geom_violin(alpha = 0.5)+
  geom_jitter(width = .05,
    size = 2,
    alpha = 0.5) +
  
  scale_fill_manual(values = fig1_palette) +

  labs(
    x = "Watershed",
    y = "LOE Temp (°C)"
  ) +

  theme_bw() +
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text = element_blank(),
    axis.title = element_text(size = 10),
    strip.text = element_text(size = 10),
    panel.spacing = unit(0.4, "lines"),
    legend.position = "bottom",
    legend.spacing.y = unit(0.05, "cm"),
  legend.margin = margin(t = 0, b = 0, unit = "cm"),
  legend.box.margin = margin(0, 0, 0, 0)
  )+
  guides(
  color = guide_legend(nrow = 2, byrow = TRUE,
                       keywidth = 0.4,
    keyheight = 0.4)
);p

sig_wsy_gh <- wsy_gh %>% 
  filter(!p.adj.signif == "ns") %>% 
  mutate(y.position = c(32, 31.9, 32.1, 32.2,32.3,32.4,32.5))

sigplot_ALL <- p+
  stat_pvalue_manual(
    sig_wsy_gh,
    label = "p.adj.signif",
    y.position = "y.position",
    inherit.aes = FALSE
  ); sigplot_ALL

ggsave("output/sigplot_ALL.png", plot = sigplot_ALL, height = 4, width = 5.25, units = "in")
```

## Basic Data Analysis of Length

### Normality check

To fit a linear model with length, I need to test the four assumptions of general linear models.

#### Does size distribution differ between locations?

I could test for normality, but Welch's ANOVA and GH post hoc are fairly robust to that.

```{r}
oneway.test(Length_mm ~ WatershedYear, data = treat2, var.equal = FALSE)
gh_length<- games_howell_test(treat2, Length_mm ~ WatershedYear); gh_length


```

```{r}
p2 <- treat2 %>% 
  ggplot(aes(x = as.factor(WatershedYear), y = Length_mm, fill = Watershed, shape = Year)) +

  geom_violin(alpha = 0.7)+
  geom_jitter(width = .05,
    size = 2,
    alpha = 0.5) +
  
  scale_fill_manual(values = fig1_palette) +

  labs(
    x = "Watershed",
    y = "Lenght (mm)"
  ) +

  theme_bw() +
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.text = element_blank(),
    axis.title = element_text(size = 10),
    strip.text = element_text(size = 10),
    panel.spacing = unit(0.4, "lines"),
    legend.position = "bottom",
    legend.spacing.y = unit(0.05, "cm"),
  legend.margin = margin(t = 0, b = 0, unit = "cm"),
  legend.box.margin = margin(0, 0, 0, 0)
  )+
  guides(
  color = guide_legend(nrow = 2, byrow = TRUE,
                       keywidth = 0.4,
    keyheight = 0.4)
);p2

sig_length <- gh_length %>% 
  filter(!p.adj.signif == "ns") %>% 
  mutate(y.position = c(215,225,235,245,255, 265))

sigplot_length <- p2+
  stat_pvalue_manual(
    sig_length,
    label = "p.adj.signif",
    y.position = "y.position",
    inherit.aes = FALSE
  ); sigplot_length

ggsave("output/sigplot_length.png", plot = sigplot_ALL, height = 4, width = 5.25, units = "in")
```

#### Linear model for effect of size and location on CTmax

```{r}
length_mod <- lm(LOE.Temp_C ~ Length_mm, data = treat2); summary(length_mod)

add_mod <- lm(LOE.Temp_C ~ Length_mm + WatershedYear, data = treat2); summary(add_mod)

int_mod <- lm(LOE.Temp_C ~ Length_mm*WatershedYear, data = treat2); summary(int_mod)
```

The current plan is to upload the stream temperature data so that it is in usable form, then use that to build the best model. After that, I can check the data for normality to see if a linear model is a good fit. If not, I will use a generalized linear model.

------------------------------------------------------------------------

## Stream Temperature

I want to maintain as much of the stream temperature data as possible in the 1-2 weeks ahead of experiments. It's important to note that the limiting factor is Big Creek in Big Sur in 2024, in which we only have a week of data, which is a limitation of the data set. We have no reason to believe that the stream temperature would have deviated from usual at that site preceding the experiments. Additionally, the Carmel River SHW site had a temp logger placed by MPWMD that blew out in 2023 and 2024, so there is a gap in the historic data set.

The first step will be recombining the data that I have into a new summary data file. I already have one from a previous script uploaded, but I want to expand the data set for Scott Creek in 2024 and Carmel River 2025.

```{r}
carmelriver <- read.csv(here("data/Stream temperature/carmel.mpwmd.Aug2024_Sept2025.csv"))
scottcreek24 <- read.csv(here("data/Stream temperature/BigCreek.ScottCreek.StreamTemp.2024-09-03.csv"))
scottcreek25 <- read.csv(here("data/Stream temperature/ScottCreekBigCreekEfish_Temp_Paige_20250520-20251014.csv"))
bigcreek24 <- read.csv(here("data/Stream temperature/BigCreek.8.26.24.csv"))
bigcreek25 <- read.csv(here("data/Stream temperature/22095921_GAUGE_POOL_12DEC2025.csv"))
```

#### First, make a list of dataframes

this part is a carryover from another script- not sure why I need to do this yet.

```{r}
#df_list <- list(bcst24,bcst25,crst24,crst25,scst24,scst25,east,dcst)
```

#### A) Clean the column names so that the temperature column is always Temp_C and the date and time column is Date_Time

#### B) Get rid of unnecessary columns

#### C) Restructure the Date_Time column

```{r}
# something is up with Scott Creek 2024 data format. Going to restructure it and try to fix it. 
```

```{r}
complete_function <- function(df, location) {
  df %>% 
    # rename columns
    dplyr::rename(
      Temp_C = dplyr::any_of(c("Temperature...C.","Temperature....C","Temperature_C","Temperature..C.")), # Rename the temperature column
      Date_Time = dplyr::any_of(c("Date.Time..PDT.","DateTime_GMT0700","timestamp","Timestamp")) # rename the Date and Time column
    ) %>% 
    # get rid of unwanted columns
    dplyr::select(-any_of(c("X.","X","Light_lux","Light..lux.","Light...lux"))
                  ) %>% 
    # Restructure the Date_Time column and add separate Date and Time columns
    mutate(
      Date_Time = lubridate::parse_date_time(
        trimws(Date_Time),
        orders = c(  "mdY HMS",
    "mdY HM",
    "ymd HMS",
    "ymd HM")
      ),
      Date = as.Date(Date_Time),
      Time = hms::as_hms(Date_Time), #standardize as hh:mm to avoid errors
      Year = as.factor(year(Date_Time)), # add a Year column
      Location = location, # when applying this, write the location name in "" within the function call 
           ) 
}
```

#### 3. Apply functions

```{r}
bcst24_cleaned <- complete_function(bigcreek24, "Big Creek") 
bcst25_cleaned <- complete_function(bigcreek25, "Big Creek")
scst24_cleaned <- complete_function(scottcreek24, "Scott Creek")
#scott creek 24 still isn't parsing correctly with this function. but it has the dates I need. So for the sake of time, I 'm not going to fix it right now but will circle back. 
scst25_cleaned <- complete_function(scottcreek25, "Scott Creek")
```

Carmel River needs a little extra cleaning! The Date and Time columns are already separate, and the temperature is in fahrenheit. The next steps:

a\) unite Date and Time columns to Date_Time and check the format

b\) convert the temp to Celsius. Get rid of site and Fahrenheit columns.

```{r}
# Cleaning up the Carmel River data frame
#a) unite date and time columns to Date_Time
carmelriver_clean <- carmelriver %>% 
  unite(Date_Time, Date, Time, sep = " ", remove= FALSE) %>%
  # b) convert F to C
  mutate(Temp_C= (Temp..F.-32)*5/9) %>% 
  # get rid of fahrenheit and site columns
  dplyr::select(-c(Temp..F., Site,Date,Time))

cr_cleaned = carmelriver_clean %>% 
  mutate(
      Date_Time = mdy_hm(Date_Time))
cr_cleaned = cr_cleaned %>% 
  mutate(Date = as.Date(Date_Time),
         Time = hms::as_hms(Date_Time), #standardize as hh:mm to avoid errors
      Year = as.factor(year(Date_Time)))


```

#### 3b. Check data frames and clean as needed

##### NOTE: get rid of first 58 rows of Big Sur 2025 data

```{r}
bcst25_cleaned <- bcst25_cleaned[-c(1:58),]
```

We also only need Scott Creek 2025 until Aug 20 and Big Sur 2025 through Aug 1.

```{r}
bcst25_cleaned <- bcst25_cleaned %>% filter(Date <= "2025-08-01")

scst25_cleaned <- scst25_cleaned %>% filter(Date <= "2025-08-20")
```

I need to separate Carmel River into two data frames: one that ends at 10/30/2024, and one that continues past that.

Maybe I should use only June temps from Carmel River in 2025

```{r}
cr24_cleaned <- cr_cleaned %>% 
  filter(Date < as.Date("2024-10-31"))
# filter just June-Aug 27
cr25_cleaned <- cr_cleaned %>% 
  filter(between(as.Date(Date), as.Date("2025-06-01"),as.Date("2025-08-27") ))
```

\

1.  Filter 8 days before experiment
2.  Plot
3.  Summary stats: 7DADM; 7DADA; 7D diel range
4.  Plot

#### Filter data 8 days before experiments

I want to calculate seven daily average daily maximum, average, and diel range. Because I only have Big Creek in 2024 through 6 days, this will be slightly abbreviated.

```{r}
# first, filter each data frame to the required dates

# Big Creek 2024: 08-20 - 08-26
# this data set is annoyingly truncated. Look to see if there is more data!!
bcst24_7d <- bcst24_cleaned %>% 
  filter(between(as.Date(Date), as.Date("2024-08-20"),as.Date("2024-08-26") )) %>% 
  group_by(Date) %>% 
  summarize(daily_max = max(Temp_C),
         daily_avg = mean(Temp_C),
         daily_min = min(Temp_C)) %>% 
  mutate(daily_diel = daily_max-daily_min,
         Watershed = "Big Creek",
         WatershedYear = "Big Creek 2024") %>% 
  ungroup()
  

# Big Creek 2025
str(bcst25_cleaned)
bcst25_7d <- bcst25_cleaned %>% 
  filter(between(as.Date(Date), as.Date("2025-07-24"),as.Date("2025-07-30") )) %>% 
  group_by(Date) %>% 
  summarize(daily_max = max(Temp_C),
         daily_avg = mean(Temp_C),
         daily_min = min(Temp_C)) %>% 
  mutate(daily_diel = daily_max-daily_min,
         Watershed = "Big Creek",
         WatershedYear = "Big Creek 2025") %>% 
  ungroup()

# Scott Creek 2024 experiments conducted9/5/24
scst24_7d <- scst24_cleaned %>% 
  filter(between(as.Date(Date), as.Date("2024-08-28"),as.Date("2024-09-03") )) %>% 
  group_by(Date) %>% 
  summarize(daily_max = max(Temp_C),
         daily_avg = mean(Temp_C),
         daily_min = min(Temp_C)) %>% 
  mutate(daily_diel = daily_max-daily_min,
         Watershed = "Scott Creek",
         WatershedYear = "Scott Creek 2024") %>% 
  ungroup()

# Scott Creek 2025: experiments conducted 8/20/2025
scst25_7d <- scst25_cleaned %>% 
  filter(between(as.Date(Date), as.Date("2025-08-12"),as.Date("2025-08-18") )) %>% 
  group_by(Date) %>% 
  summarize(daily_max = max(Temp_C),
         daily_avg = mean(Temp_C),
         daily_min = min(Temp_C)) %>% 
  mutate(daily_diel = daily_max-daily_min,
         Watershed = "Scott Creek",
         WatershedYear = "Scott Creek 2025") %>% 
  ungroup()

# Carmel River 2024 experiments conducted 10/30/24
colnames(cr24_cleaned)
crst24_7d <- cr24_cleaned %>% 
  filter(between(as.Date(Date), as.Date("2024-10-22"),as.Date("2024-10-28") )) %>% 
  group_by(Date) %>% 
  summarize(daily_max = max(Temp_C),
         daily_avg = mean(Temp_C),
         daily_min = min(Temp_C)) %>% 
  mutate(daily_diel = daily_max-daily_min,
         Watershed = "Carmel River",
         WatershedYear = "Carmel River 2024") %>% 
  ungroup()

# Carmel River 2025
crst25_7d <- cr25_cleaned %>% 
  filter(between(as.Date(Date), as.Date("2025-08-19"),as.Date("2025-08-25") )) %>% 
  group_by(Date) %>% 
  summarize(daily_max = max(Temp_C),
         daily_avg = mean(Temp_C),
         daily_min = min(Temp_C)) %>% 
  mutate(daily_diel = daily_max-daily_min,
         Watershed = "Carmel River",
         WatershedYear = "Carmel River 2025") %>% 
  ungroup()

```

```{r}
# Combine the data frames into one
stream7d = rbind(bcst24_7d, bcst25_7d, scst24_7d, scst25_7d, crst24_7d,crst25_7d)
# add a column count for total days
```

#### Plot 7d stream temperature

```{r}
colnames(stream7d)
ggplot(stream7d, aes(x = Date, y = daily_avg, color = Watershed))+
  geom_line()+
  facet_wrap(~WatershedYear)
```

## Build model

This model will use a glmm framework with the goal of building an effect size plot.

Don't forget to check for correlation between variables using ggpairs!\

```{r}
stream7d_sum = stream7d %>% 
  group_by(WatershedYear) %>% 
  summarize(DADM7 = mean(daily_max),
            sd_DADM7 = sd(daily_max),
            DADA7 = mean(daily_avg),
            sd_DADA7 = sd(daily_avg),
            DADMin7 = mean(daily_min),
            DADDiel7 = mean(daily_diel))%>% 
  ungroup()

# save the data frame for other uses
write.csv(stream7d_sum, "data/stream7d_sum.ch1.csv", row.names = FALSE)
```

Leaving off on 5/28:

-   add a column for number of days (n 1-7) to visualize plot better. Maybe add text with the date ranges for each site?

-   DONE: make a new data frame that summarizes into one value that can be inputted into model

-   add in historical data? (might as well!) - or just add in effect of location

-   maybe location can be both a fixed and random effect in the model?

-   ggpairs to look for correlation between variables

### Look for autocorrelation between variables

First, I need to combine data frames!

```{r}
# group by WatershedYear
treat2_st = full_join(treat2, stream7d_sum, by = "WatershedYear")
```

Next, check for autocorrelation

```{r}
library(GGally)
ggpairs(treat2_st, columns = c("Length_mm","DADM7","DADA7","Watershed","DADDiel7"))
```

From this ggpairs output, I should use only one stream temperature output at a time in my model.

NEXT: Do I need to put in correlated variables from LOE.Temp_C and WatershedYear?

## Model Design

```{r}
model = lmer(LOE.Temp_C ~ Length_mm + DADM7 + Watershed + (1|WatershedYear), 
             data = treat2_st); summary(model)
```
