---
title: "ThermalPhysiologyFigures"
format: html
editor: visual
---

## Figures and Exploratory Analysis

This code is intended to set the stage for further analyses.

1.  Load in data and clean it up

2.  Make preliminary plots:

    -   fish length distribution

    -   Plots for visualizing CTmax in 2024 vs 2025 for just treatment 2

    -   Stream temperature

3.  Basic data analysis:

    -   Check data for normality (probably going to need to do a Welch's ANOVA with Games-Howell posthoc test for pairwise comparisons)

    -   Is size distribution within locations different across years?

    -   Is size distribution across locations different within years?

    -   Is CTmax different within locations across years?

    -   Is CTmax different across locations within years?

## Part 1: Load in Data and Clean

### Libraries

```{r}
lapply(c("tidyverse", "lme4", "ggthemes","car","emmeans","brglm2", "ggthemes", 
         "rethinking", "usethis", "here", "mgcv", "gratia", "extrafont"), 
       require, character.only=T)
```

### Data

#### Physiology

```{r}
comp <- read.csv(here("data/IRA.thermaltolerance.compileddata.csv"))
head(comp)
```

##### 1. Rename columns and check structure

I am going to rename the Location column to Watershed, since there are two Big Creek's. This will make it more straightforward later on with color palettes and clarity in figures.

```{r}
comp_clean = comp %>% 
  rename(Watershed = Location)
unique(comp_clean$Watershed)

# rename locations to match the watershed name
comp_clean = comp_clean %>% 
    mutate(Watershed = case_when(
    Watershed == "Big Sur" ~ "Big Creek", # Big Creek in Big Sur
    Watershed == "Napa" ~ "Napa River", # Napa River
    Watershed == "E Austin" ~ "Russian River", # Russian River
    TRUE~ Watershed # keep Scott Creek and Carmel River the same
  ))
```

```{r}
str(comp_clean)
comp_clean$Date = as.Date(comp$Date, format = "%m/%d/%Y") #convert date to date structure
# Treatment should be a factor
comp_clean$Treatment = as.factor(comp_clean$Treatment)

# check for NA values
sum(is.na(comp_clean))
# there is one NA value in the length column

# add a year column
comp_clean <- comp_clean %>% 
  mutate(Year = year(Date))
comp_clean$Year = as.factor(comp_clean$Year)
comp_clean$Mass_g <- as.numeric(comp_clean$Mass_g)

# Add a LocationYear column
comp_clean <- comp_clean %>% 
  unite(WatershedYear, Watershed, Year, sep = " ", remove= FALSE)

# add a sample size column
#comp %>% 
#  group_by(Location) %>% 
#  summarise(sample_size = n())

# calculate relative weight
comp_clean <- comp_clean %>%
  mutate(
    #Where are these weight-length parameters from? FishBase?
    log10Ws = -4.790 + 2.928 * log10(Length_mm), 
    #Why taking log here
    Ws = 10^log10Ws
  )

comp_clean <- comp_clean %>%
  mutate(
    Wr = (Mass_g / Ws)*100
  )

```

#### Treatment 2 data ONLY

This is the dataframe we are tentatively planning on using for just chapter 1. This will be more or less a purely physiological paper.

From Big Creek, Scott Creek, and Carmel River

```{r}
treat2 = comp_clean %>% 
  filter(Treatment == 2) %>% 
  filter(Watershed %in% c("Big Creek", "Carmel River", "Scott Creek"))
```

#### Stream Temperature

### Color Palettes

```{r}
# palette generated by Coolors
fig1_palette= c(
  "Big Creek"="#744253",
  "Scott Creek"="#496A81",
  "Carmel River"="#9BC53D")
```

## Part 2: Preliminary Plots

Because this paper will only use Big Creek, Scott Creek, and Carmel River, these plots will only focus on those three sites.

### Size distribution

```{r}
size_dist <- ggplot(treat2, aes(x = Length_mm, fill = Watershed))+
  geom_histogram(binwidth = 5, , color = "black")+
  scale_fill_manual(values = fig1_palette)+
   scale_y_continuous(
    expand = expansion(mult = c(0, 0))
  )+
  labs(x = "Length (mm)",
       y = "Count")+
  theme_bw()+
  facet_grid(Watershed~Year)

size_dist
```

#### Within locations, across years

```{r}
size_distbyyear <- ggplot(treat2, aes(x = Length_mm, fill = Watershed))+
  geom_histogram(binwidth = 5, , color = "black")+
  scale_fill_manual(values = fig1_palette)+
   scale_y_continuous(
    expand = expansion(mult = c(0, 0))
  )+
  labs(x = "Length (mm)",
       y = "Count")+
  theme_bw()+
  facet_wrap(~Year)

size_distbyyear
```

```{r}
size_distbylocation <- ggplot(treat2, aes(x = Length_mm, fill = Watershed))+
  geom_histogram(binwidth = 5, , color = "black")+
  scale_fill_manual(values = fig1_palette)+
   scale_y_continuous(
    expand = expansion(mult = c(0, 0))
  )+
  labs(x = "Length (mm)",
       y = "Count")+
  theme_bw()+
  facet_wrap(~Watershed)

size_distbylocation
```

### Critical Thermal Maximum

```{r}
treat2 %>% 
  group_by(WatershedYear) %>% 
  mutate(mean_C = mean(LOE.Temp_C)) %>% 
  ungroup() %>% 
  filter(!Treatment == 1) %>% 
  ggplot(aes(x = as.factor(Watershed), y = LOE.Temp_C, color = Watershed)) +

  geom_jitter(width = .25,
    size = 2.5,
    alpha = 0.4) +

  stat_summary(
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.8),
    size = 4,
    stroke = 0.6
  ) +

  facet_wrap(~Year) +

  scale_color_manual(values = fig1_palette) +

  labs(
    x = "Watershed",
    y = "LOE Temp (°C)"
  ) +

  theme_bw() +
  theme(
    text = element_text(family = "Times New Roman", size = 8),
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 7),
    strip.text = element_text(size = 7),
    panel.spacing = unit(0.4, "lines"),
    legend.position = "bottom",
    legend.spacing.y = unit(0.05, "cm"),
  legend.margin = margin(t = 0, b = 0, unit = "cm"),
  legend.box.margin = margin(0, 0, 0, 0)
  )+
  guides(
  color = guide_legend(nrow = 2, byrow = TRUE,
                       keywidth = 0.4,
    keyheight = 0.4)
)

# may need to make points smaller for when I save the figure for the manuscript
```

### Stream Temperature

## Basic Data Analysis
