---
title: "Lit Comparison"
format: html
editor: visual
---

## Literature Comparion

To compare our measured upper thermal limits to those measured in other populations, I pulled literature that followed similar methods. The selected papers had to meet the following criteria:

-   Use the same ramp rate (0.3°C per min)

-   Juvenile (age 0-1; or 50mm-200mm FL)

-   take place on the West coast of North America

-   Use a similar acclimation/holding temperature (20°C +/- 2°C)

I did not consider the following:

-   **Holding duration:** while we know it matters (citation), the 20°C holding duration employed in our study and 2 other represented here is not a common enough method to narrow our search to this requirement.

-   **Hatchery vs wild**: streamside thermal tolerance experiments are logistically challenging and don't always meet a study's needs; again, this would narrow the search too extensively to produce meaningful results

-   **Rainbow trout vs steelhead:** there seems to be little to no difference between these two groups of *O. mykiss*. However, we exclude subspecies (i.e., redband trout).

The following papers are included:

-   Myrick, Christopher A., and Joseph J. Cech Jr. “Effects of Temperature on the Growth, Food Consumption, and Thermal Tolerance of Age-0 Nimbus-Strain Steelhead.” *North American Journal of Aquaculture* 67, no. 4 (2005): 324–30. <https://doi.org/10.1577/A04-050.1>.

-   Dressler, Terra L., Kara Anlauf-Dunn, Andrea Chandler, and Erika J. Eliason. “Beyond Latitude: Thermal Tolerance and Vulnerability of a Broadly Distributed Salmonid across a Habitat Temperature Gradient.” *Conservation Physiology* 13, no. 1 (2025): coaf030. <https://doi.org/10.1093/conphys/coaf030>.

-   Dressler, T. L., V. Han Lee, K. Klose, and E. J. Eliason. “Thermal Tolerance and Vulnerability to Warming Differ between Populations of Wild Oncorhynchus Mykiss near the Species’ Southern Range Limit.” *Scientific Reports* 13, no. 1 (2023): 14538. <https://doi.org/10.1038/s41598-023-41173-7>.

-   Myrick, C. A., and J. J. Cech. “Temperature Influences on California Rainbow Trout Physiological Performance.” *Fish Physiology and Biochemistry* 22, no. 3 (2000): 245–54. <https://doi.org/10.1023/A:1007805322097>.

-   Recsetar, Matthew S., Matthew P. Zeigler, David L. Ward, Scott A. Bonar, and Colleen A. Caldwell. “Relationship between Fish Size and Upper Thermal Tolerance.” *Transactions of the American Fisheries Society* 141, no. 6 (2012): 1433–38. <https://doi.org/10.1080/00028487.2012.694830>.

## Plotting the comparison

### Libraries

```{r}
library(tidyverse)
library(here)
library(extrafont)
```

### Data

```{r}
lit <- read.csv(here("data/ctmaxlit_comparison.csv"))

# get rid of row 17 (no data there)
lit <- head(lit, -1)
```

### Plot

```{r}
fig_palette= c(
  "Big Creek"="orange",
  "Scott Creek"="cyan",
  "Carmel River 2024"="#9BC53D",
  "Carmel River 2025"="chartreuse4",
  "Napa River"="#F991CC",
  "Russian River" = "purple",
  "Nimbus Hatchery"="black",
  "Eagle Lake"="black", 
  "Mt. Shasta Fish Hatchery"="black", 
  "John Day River"="black",           
  "Deschutes River"="black",         
  "Siletz River"="black",             
  "Umpqua River" ="black",            
  "Piru Creek"="black",               
  "Arroyo Seco Creek" ="black",       
  "Arizona Hatchery"="black"   
  )
colnames(lit)
lit$Watershed
```

```{r}
lit_plot <- ggplot(lit, aes(x = RearingTemp_C, y = CTmax, color=Watershed)) +
  geom_point(size = 3) +
  scale_color_manual(values=fig_palette)+
  labs(
    x = "Maximum Rearing Temperature (°C)",
    y = "Loss of Equilibrium Temperature (°C)"
  ) +
  theme_bw()+
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size=10),
    axis.text.y = element_text(size=10),
    strip.text = element_text(size = 10),
    panel.spacing = unit(0.4, "lines"),
    legend.position = "none"
  );lit_plot
  

```

```{r}
ggsave("output/lit_comparison.png", plot = lit_plot, height = 4, width = 4, dpi= 300, units = "in")
```
