---
title: "Results figures"
format: html
editor: visual
---

## Chapter 1: Results Figures

The goal of this script is to present all the relevant analyses and figures to be included in the Chapter 1 manuscript. Chapter 1 seeks to answer the following questions:

1.  What is the critical thermal maximum of five (three?) coastal CA populations in response to a common, ecologically relevant acute temperature treatment?

2.  Does age/size class for juvenile steelhead impact upper thermal limits? (will increasingly variable temps disproportionately impact size classes)

3.  Is there evidence for the role of plasticity and local adaptation in determining upper thermal limits?

This document follows the order of the above questions.

### Libraries

```{r}
library(ggplot2)
library(ggthemes)
library(tidyverse)
library(lme4)
library(here)
library(extrafont)
library(RColorBrewer)
```

### Data

```{r}
# Load in data --------------------------------------------------------
comp <- read.csv(here("data/IRA.thermaltolerance.compileddata.csv"))
head(comp)

## Wrangling -----------------------------------------------------------
comp_clean = comp %>% 
  rename(Watershed = Location)
unique(comp_clean$Watershed)

# rename locations to match the watershed name
comp_clean = comp_clean %>% 
  mutate(Watershed = case_when(
    Watershed == "Big Sur" ~ "Big Creek", # Big Creek in Big Sur
    Watershed == "Napa" ~ "Napa River", # Napa River
    Watershed == "E Austin" ~ "Russian River", # Russian River
    TRUE~ Watershed # keep Scott Creek and Carmel River the same
  ))
str(comp_clean)

comp_clean$Date = as.Date(comp$Date, format = "%m/%d/%Y") #convert date to date structure
# Treatment should be a factor
comp_clean$Treatment = as.factor(comp_clean$Treatment)

# check for NA values
sum(is.na(comp_clean))
# there is one NA value in the length column

# add a year column
comp_clean <- comp_clean %>% 
  mutate(Year = year(Date))
comp_clean$Year = as.factor(comp_clean$Year)
comp_clean$Mass_g <- as.numeric(comp_clean$Mass_g)

# Add a LocationYear column
comp_clean <- comp_clean %>% 
  unite(WatershedYear, Watershed, Year, sep = " ", remove= FALSE)

## Filtered dataset ----------------------------------------------------------------------------------------------
treat2 = comp_clean %>% 
  filter(Treatment == 2) %>% 
  filter(Watershed %in% c("Big Creek", "Carmel River", "Scott Creek"))

treat2ALL = comp_clean %>% 
  filter(Treatment == 2)

```

### Color Palettes

```{r}
fig1_palette= c(
  "Big Creek"="#744253",
  "Scott Creek"="#496A81",
  "Carmel River"="#9BC53D")

fig2_palette = c(
  "Big Creek 2024"="#C89F9C",
  "Big Creek 2025" = "#B36A5E",
  "Scott Creek 2025"="#083d77",
  "Scott Creek 2024"="#76949f",
  "Carmel River 2024" = "#9BC53D",
  "Carmel River 2025" = "chartreuse4",
  "Russian River 2025" = "purple",
  "Napa River 2025" = "orange"
)

fig3_palette= c(
  "Big Creek"="darkorange",
  "Scott Creek"="cyan4",
  "Carmel River"="chartreuse4",
  "Napa River"="#F991CC",
  "Russian River" = "purple")

fig4_palette= c(
  "Big Creek 2024"="orange",
  "Big Creek 2025"="darkorange2",
  "Scott Creek 2024"="cyan",
  "Scott Creek 2025"="cyan4",
  "Carmel River 2024"="#9BC53D",
  "Carmel River 2025"="chartreuse4",
  "Napa River 2025"="#F991CC",
  "Russian River 2025" = "purple")
```

## Part 1: Quantifying CTmax for 5 CA steelhead populations

First, I will make a figure showing CTmax for each site and year.

```{r}
# set levels according to latitude
# 1. Define your explicit order
lat_order <- c("Big Creek", "Carmel River", "Scott Creek","Napa River","Russian River")

# 2. Re-level the data frame column
treat2ALL$Watershed <- factor(treat2ALL$Watershed, levels = lat_order)


# lat and year order
latyear_order <- c("Big Creek 2024", "Big Creek 2025","Carmel River 2024", "Carmel River 2025", "Scott Creek 2024","Scott Creek 2025","Napa River 2025","Russian River 2025")
treat2ALL$WatershedYear <- factor(treat2ALL$WatershedYear, levels = latyear_order)
```

```{r}

RColorBrewer::display.brewer.all()

p <- treat2ALL %>% 
  ggplot(aes(x = as.factor(Year), 
             y = LOE.Temp_C, fill = WatershedYear)) +
  
  geom_violin(alpha = 0.75)+
  geom_jitter(width = .05,
              size = 2.5,
              alpha = 0.75) +
  
  scale_fill_manual(values= fig4_palette) +
  
  labs(
    x = "Sample Year",
    y = "LOE Temp (°C)"
  ) +
  
  theme_bw() +
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size=10),
    axis.text.y = element_text(size=10),
    strip.text = element_text(size = 10),
    panel.spacing = unit(0.4, "lines"),
    legend.position = "none"
  )+
  guides(
    color = guide_legend(nrow = 2, byrow = TRUE,
                         keywidth = 0.4,
                         keyheight = 0.4)
  )+
  facet_grid(~Watershed);p

ggsave("output/treat2ALLctmax.png", plot = p, height = 5, width = 6, dpi= 300, units = "in")
```

## Part 2: Length effect

I assumed the effect of length on CTmax would be conserved across populations and not vary intraspecifically. This is because there is no evidence to suggest otherwise, and also to increase sample size.

```{r}
ggplot(treat2ALL, aes(x= Length_mm, y = LOE.Temp_C))+
  geom_point(size=2,
             alpha = 0.75)+
  geom_smooth(method= lm)+
  labs(x= "Critical Thermal Maximum (°C)",
       y= "Fork Length (mm)")+
  theme_bw()+
  theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size=10),
    axis.text.y = element_text(size=10),
    strip.text = element_text(size = 10),
    panel.spacing = unit(0.4, "lines"),
    legend.position = "none"
  )
```

Check for normality- might need to a generalized linear model!

```{r}
fit = lm(LOE.Temp_C~Length_mm, data=treat2ALL)

# checks
par(mfrow=c(2,2))
plot(fit)

#Normalityofresiduals:
shapiro.test(resid(fit))
# p > 0.05 (barely) - seems to be fairly normal

#Linearity of the relationship:
summary(lm(resid(fit)~poly(predict(fit),2)))
# the second term (,2) is not significant, so linear enough

#Homoscedasticity (constant variance):
library(lmtest)
bptest(fit)
# slightly heteroskedastic (p = 0.04)

#Independence of residuals for temporal data
#(i.e.data points taken on successive days/weeks/etc.);
#your response variable needs to be sorted by date:
pacf(resid(fit))

# this is showing that individuals more close in size are more likely to be similar in CTmax. I don't think this one is relevant, but will check
```

It seems we can assume normality, and use a linear model. There is some slight violation of homoskedasticity and independence, but for interpretation and biological purposes, we are going to be ok with that.

```{r}
summary(fit)
confint(fit, level=0.95) # extracet 95% CI's

```

The effects is small but significant (p\<0.05, \*\*\*). With every mm increase in fork length, CTmax decreases by -0.0031°C (CI ∓ 0.001596). Length explains \~8% of the variation in CTmax.

```{r}
library(ggpubr)
sizeeffect<- ggplot(data=treat2ALL,aes(x=Length_mm,y=LOE.Temp_C, color=Watershed))+
  geom_point(size= 2)+
  geom_smooth(method="lm")+
  stat_cor(label.y = c(32.15),color="purple")+
  stat_regline_equation(label.y = c(32),color="purple")+
  labs(x= "Fork length (mm)",
       y = "Critical Thermal Maximum (°C)")+
  theme_bw()+
   theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size=10),
    axis.text.y = element_text(size=10),
    strip.text = element_text(size = 10),
    panel.spacing = unit(0.4, "lines")#,
   # legend.position = "none"
  ); sizeeffect

#ggsave("output/sizeeffect.png", plot = sizeeffect, height = 5, width = 6, dpi= 300, units = "in")
  
```

Now we just need a raw plot of the size distributions! Let's try ggridges and a bar plot

### Size distribution

```{r}
## GGridges
library(ggridges)

ggridges_size<- ggplot(treat2ALL, aes(x= Length_mm, y= Watershed, fill = WatershedYear))+
  geom_density_ridges(alpha = 0.75)+
  scale_fill_manual(values= fig4_palette)+
  #scale_fill_brewer(palette = "Paired") +
  labs(x="Length (mm)",
       y = "",
       fill= "Site/Year")+
  theme_bw()+
    theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.title = element_text(size = 10),
    axis.title.x = element_text(hjust = 0.5),
    axis.text.x = element_text(size=10),
    axis.text.y = element_text(size=10),
    strip.text = element_text(size = 10),
    panel.spacing = unit(0.4, "lines")
  ); ggridges_size
ggsave("output/ggridges_size.png", plot = ggridges_size, height = 3.5, width = 6, units = "in", dpi= 300)
```

## Part 3: Evidence for the role of plasticity and local adaptation in upper thermal limits

-   GLM: Effect of fish size, 7DADM, and Watershed (both a random and fixed effect)

    -   7DADM is a management relevant metric

    -   Fish size generally seems to decrease thermal tolerance

### Load in stream temp data

```{r}
# Load in data
carmelriver <- read.csv(here("data/Stream temperature/carmel.mpwmd.Aug2024_Sept2025.csv"))
scottcreek24 <- read.csv(here("data/Stream temperature/BigCreek.ScottCreek.StreamTemp.2024-09-03.csv"))
scottcreek25 <- read.csv(here("data/Stream temperature/ScottCreekBigCreekEfish_Temp_Paige_20250520-20251014.csv"))
bigcreek24 <- read.csv(here("data/Stream temperature/BigCreek.8.26.24.csv"))
bigcreek25 <- read.csv(here("data/Stream temperature/22095921_GAUGE_POOL_12DEC2025.csv"))
eaustin25 <- read.csv(here("data/Stream temperature/eastaustin.streamtemphobo.may_sept2025.csv"))
napa25 <- read.csv(here("data/Stream temperature/Dry creek Napa  2026-06-17 13_24_54 PDT.csv"))

complete_function <- function(df, location) {
  df %>% 
    # rename columns
    dplyr::rename(
      Temp_C = dplyr::any_of(c("Temperature...C.","Temperature....C","Temperature_C","Temperature..C.","Temperature.....C.")), # Rename the temperature column
      Date_Time = dplyr::any_of(c("Date.Time..PDT.","DateTime_GMT0700","timestamp","Timestamp","Date.Time..PDT.PST.")) # rename the Date and Time column
    ) %>% 
    # get rid of unwanted columns
    dplyr::select(-any_of(c("X.","X","Light_lux","Light..lux.","Light...lux","Light....lux."))
                  ) %>% 
    # Restructure the Date_Time column and add separate Date and Time columns
    mutate(
      Date_Time = lubridate::parse_date_time(
        trimws(Date_Time),
        orders = c(  "mdY HMS",
    "mdY HM",
    "ymd HMS",
    "ymd HM")
      ),
      Date = as.Date(Date_Time),
      Time = hms::as_hms(Date_Time), #standardize as hh:mm to avoid errors
      Year = as.factor(year(Date_Time)), # add a Year column
      Location = location, # when applying this, write the location name in "" within the function call 
           ) 
}


bcst24_cleaned <- complete_function(bigcreek24, "Big Creek") 
bcst25_cleaned <- complete_function(bigcreek25, "Big Creek")
scst24_cleaned <- complete_function(scottcreek24, "Scott Creek")
scst25_cleaned <- complete_function(scottcreek25, "Scott Creek")
dc25_cleaned <- complete_function(napa25, "Napa River")
dc25_cleaned <- dc25_cleaned %>% 
  dplyr::select(c("Date_Time","Temp_C","Date","Time","Year","Location"))

ea25_cleaned<- complete_function(eaustin25,"Russian River")

# Cleaning up the Carmel River data frame
#a) unite date and time columns to Date_Time
carmelriver_clean <- carmelriver %>% 
  unite(Date_Time, Date, Time, sep = " ", remove= FALSE) %>%
  # b) convert F to C
  mutate(Temp_C= (Temp..F.-32)*5/9) %>% 
  # get rid of fahrenheit and site columns
  dplyr::select(-c(Temp..F., Site,Date,Time))

cr_cleaned = carmelriver_clean %>% 
  mutate(
      Date_Time = mdy_hm(Date_Time))
cr_cleaned = cr_cleaned %>% 
  mutate(Date = as.Date(Date_Time),
         Time = hms::as_hms(Date_Time), #standardize as hh:mm to avoid errors
      Year = as.factor(year(Date_Time)))

#NOTE: get rid of first 58 rows of Big Sur 2025 data
bcst25_cleaned <- bcst25_cleaned[-c(1:58),]

#We also only need Scott Creek 2025 until Aug 20 and Big Sur 2025 through Aug 1.
bcst25_cleaned <- bcst25_cleaned %>% filter(Date <= "2025-08-01")

scst25_cleaned <- scst25_cleaned %>% filter(Date <= "2025-08-20")

# separate Carmel into two data frames
cr24_cleaned <- cr_cleaned %>% 
  filter(Date < as.Date("2024-10-31"))
# filter just June-Aug 27
cr25_cleaned <- cr_cleaned %>% 
  filter(between(as.Date(Date), as.Date("2025-06-01"),as.Date("2025-08-27") ))
```

Now we need to filter the data down to 8 days before experiments to calculate 7 day averages.

```{r}
# first, filter each data frame to the required dates

# Big Creek 2024: 08-20 - 08-26
# this data set is annoyingly truncated. Look to see if there is more data!!
bcst24_7d <- bcst24_cleaned %>% 
  filter(between(as.Date(Date), as.Date("2024-08-20"),as.Date("2024-08-26") )) %>% 
  group_by(Date) %>% 
  summarize(daily_max = max(Temp_C),
         daily_avg = mean(Temp_C),
         daily_min = min(Temp_C)) %>% 
  mutate(daily_diel = daily_max-daily_min,
         Watershed = "Big Creek",
         WatershedYear = "Big Creek 2024") %>% 
  ungroup()
  

# Big Creek 2025
str(bcst25_cleaned)
bcst25_7d <- bcst25_cleaned %>% 
  filter(between(as.Date(Date), as.Date("2025-07-24"),as.Date("2025-07-30") )) %>% 
  group_by(Date) %>% 
  summarize(daily_max = max(Temp_C),
         daily_avg = mean(Temp_C),
         daily_min = min(Temp_C)) %>% 
  mutate(daily_diel = daily_max-daily_min,
         Watershed = "Big Creek",
         WatershedYear = "Big Creek 2025") %>% 
  ungroup()

# Scott Creek 2024 experiments conducted9/5/24
scst24_7d <- scst24_cleaned %>% 
  filter(between(as.Date(Date), as.Date("2024-08-28"),as.Date("2024-09-03") )) %>% 
  group_by(Date) %>% 
  summarize(daily_max = max(Temp_C),
         daily_avg = mean(Temp_C),
         daily_min = min(Temp_C)) %>% 
  mutate(daily_diel = daily_max-daily_min,
         Watershed = "Scott Creek",
         WatershedYear = "Scott Creek 2024") %>% 
  ungroup()

# Scott Creek 2025: experiments conducted 8/20/2025
scst25_7d <- scst25_cleaned %>% 
  filter(between(as.Date(Date), as.Date("2025-08-12"),as.Date("2025-08-18") )) %>% 
  group_by(Date) %>% 
  summarize(daily_max = max(Temp_C),
         daily_avg = mean(Temp_C),
         daily_min = min(Temp_C)) %>% 
  mutate(daily_diel = daily_max-daily_min,
         Watershed = "Scott Creek",
         WatershedYear = "Scott Creek 2025") %>% 
  ungroup()

# Carmel River 2024 experiments conducted 10/30/24
colnames(cr24_cleaned)
crst24_7d <- cr24_cleaned %>% 
  filter(between(as.Date(Date), as.Date("2024-10-22"),as.Date("2024-10-28") )) %>% 
  group_by(Date) %>% 
  summarize(daily_max = max(Temp_C),
         daily_avg = mean(Temp_C),
         daily_min = min(Temp_C)) %>% 
  mutate(daily_diel = daily_max-daily_min,
         Watershed = "Carmel River",
         WatershedYear = "Carmel River 2024") %>% 
  ungroup()

# Carmel River 2025
crst25_7d <- cr25_cleaned %>% 
  filter(between(as.Date(Date), as.Date("2025-08-19"),as.Date("2025-08-25") )) %>% 
  group_by(Date) %>% 
  summarize(daily_max = max(Temp_C),
         daily_avg = mean(Temp_C),
         daily_min = min(Temp_C)) %>% 
  mutate(daily_diel = daily_max-daily_min,
         Watershed = "Carmel River",
         WatershedYear = "Carmel River 2025") %>% 
  ungroup()

# Napa River 2025
dcst25_7d <- dc25_cleaned %>% 
  filter(between(as.Date(Date), as.Date("2025-07-17"),as.Date("2025-07-23") )) %>% 
  group_by(Date) %>% 
  filter(!is.na(Temp_C)) %>% 
  summarize(daily_max = max(Temp_C),
         daily_avg = mean(Temp_C),
         daily_min = min(Temp_C)) %>% 
  mutate(daily_diel = daily_max-daily_min,
         Watershed = "Napa River",
         WatershedYear = "Napa River 2025") %>% 
  ungroup()

# Russian River 2025
east25_7d <- ea25_cleaned %>% 
    filter(between(as.Date(Date), as.Date("2025-09-03"),as.Date("2025-09-9") )) %>% 
  group_by(Date) %>% 
  filter(!is.na(Temp_C)) %>% 
  summarize(daily_max = max(Temp_C),
         daily_avg = mean(Temp_C),
         daily_min = min(Temp_C)) %>% 
  mutate(daily_diel = daily_max-daily_min,
         Watershed = "Russian River",
         WatershedYear = "Russian River 2025") %>% 
  ungroup()

# Combine!
stream7d = rbind(bcst24_7d, bcst25_7d, scst24_7d, scst25_7d, crst24_7d,crst25_7d, dcst25_7d,east25_7d)

stream7d_sumALL = stream7d %>% 
  group_by(WatershedYear) %>% 
  summarize(DADM7 = mean(daily_max),
            sd_DADM7 = sd(daily_max),
            DADA7 = mean(daily_avg),
            sd_DADA7 = sd(daily_avg),
            DADMin7 = mean(daily_min),
            DADDiel7 = mean(daily_diel))%>% 
  ungroup()

# save the data frame for other uses
write.csv(stream7d_sumALL, "data/stream7d_sumALL.csv", row.names = FALSE)

# Combine with treat2ALL
treat2ALL_st = full_join(treat2ALL, stream7d_sumALL, by = "WatershedYear")
```

### Models, finally!

```{r}
mod1.1 <- lm(LOE.Temp_C ~ Watershed + DADM7 + 
                Length_mm 
                 , 
               data = treat2ALL_st)
summary(mod1.1)
AIC(mod1.1)

```

The model revealed a positive, significant relationship between LOE and 7DADM, where average LOE increased at a rate of 0.10°C for every degree increase in 7DADM (95% CI 0.07-0.13, p \<0.05). However, size did not impact CTmax at all, and there was additional variation between watersheds that was unexplained by 7DADM alone. This indicates an important role of long term thermal history in plasticity that an acute holding treatment does not overcome, and also points to a possibility of local adaptation.

```{r}
mod1.2 <- lmer(LOE.Temp_C ~ DADM7 + Length_mm + (1|Watershed), data = treat2ALL_st)
summary(mod1.2)

```

```{r}
coef(mod1.1)
VarCorr(mod1.1)
resid(mod1.1)
str(mod1.1)


library(broom)

df <- model.frame(mod1.1)
df$.fitted <- predict(mod1.1)


modplot<- ggplot(df, aes(x=DADM7, y= LOE.Temp_C, color=Watershed)) +
  geom_jitter(width = 0.075, size=2,alpha = 0.6)+
  scale_color_manual(values= fig3_palette) +
  labs(x= "7 Day Average Daily Maximum (°C)",
       y = "Loss of Equilibrium Temp (°C)")+
  theme_bw()+
   theme(
    text = element_text(family = "Times New Roman", size = 10),
    axis.title = element_text(size = 10),
    axis.title.x = element_text(hjust = 0.5),
    axis.text.x = element_text(size=10),
    axis.text.y = element_text(size=10),
    strip.text = element_text(size = 10),
    panel.spacing = unit(0.4, "lines")
  ); modplot

ggsave("output/mod_output.png", plot = modplot, height = 3.5, width = 6, units = "in", dpi= 300)


```

```{r}
AIC(mod1.1)
AIC(mod1.2)
```

```{r}
mod1.3 <- lmer(LOE.Temp_C ~ Watershed + 
                 (1|Watershed), 
               data = treat2ALL_st)
summary(mod1.3)
AIC(mod1.3)

mod1.4 <- lmer(LOE.Temp_C ~ Watershed*DADM7 + 
                # Length_mm +
                 (1|Watershed), 
               data = treat2ALL_st)
summary(mod1.4)
AIC(mod1.4)
```

Mod1.1 excludes the size effect and mod1.2 includes it. However, mod1.1 has a much lower AIC, and so is a better model. Mod1.3 is the worst, and mod1.4 is also rejected.

```{r}
# check the residuals
df <- data.frame(predicted=predict(mod1.1), residual = residuals(mod1.1))
ggplot(df,aes(sample=residual)) + stat_qq() + stat_qq_line()
ggplot(df,aes(x=residual)) + geom_histogram(bins=20, color="black")
```

In writing this, report the results first of the one that I prefer. We used the model with watershed as a random effect. We also ran the model with watershed as a fixed effect and it didn't meaningfully change the output.

Justifying fixed vs random effects - if I want to know how watersheds are different,

To do:

-   the size estimate is wrong! there is no effect, shown by the lmer output.

-   plot it! use predict function

Alright so we're going to use Watershed as a fixed effect, so I can compare watersheds after adjusting for the effect of DADM7. Also, when using watershed as a random effect, the output is still the same (effect of 7DADM is the same, effect of length is the same, but now we have between watershed comparisons).

## Make a table for pub

This table should include CTmax +/- sd, 7DADM, 7DADA, fish length +/- sd, Watershed, Year

```{r}
library(gt)
library(dplyr)
library(webshot2)
class(treat2ALL_st)

gt_table <- treat2ALL_st %>% 
  group_by(Watershed, Year) %>% 
  mutate(LOE_mean = mean(LOE.Temp_C),
         LOE_sd = sd(LOE.Temp_C),
         Length_mean = mean(Length_mm),
         Length_sd = sd(Length_mm),
         date_md = format(Date, "%m-%d")) %>%
  mutate(LOE_mean= sprintf("%.2f",LOE_mean),
         LOE_sd= sprintf("%.2f",LOE_sd),
         Length_mean= sprintf("%.2f",Length_mean),
         Length_sd= sprintf("%.2f",Length_sd)) %>% 
  ungroup() %>% 
  unite(., col = "Length_mean_sd", Length_mean, Length_sd, sep = "±") %>% 
  unite(., col = "LOE_mean_sd", LOE_mean, LOE_sd, sep = "±") %>% 
  dplyr::select(Watershed, Year, date_md, DADM7, DADA7, Length_mean_sd, LOE_mean_sd) %>% 
  distinct() %>% 
  arrange(Watershed, Year) %>%
  group_by(Watershed) %>%
  mutate(Watershed = if_else(row_number() == 1, Watershed, "")) %>% 
  ungroup() %>% 
  gt() %>% 
  cols_align(align = "center") %>% 
  tab_header(
    title= "Recorded upper thermal limits"#,
  #subtitle = 
  ) %>% 
  fmt_number(
    columns = c(Watershed, Year, DADM7, DADA7, Length_mean_sd,LOE_mean_sd),
    decimals = 2
  ) %>% 
  cols_label(
    date_md = md("Date  \n (m-d)"),
    DADM7 = "7DADM (°C)",
    DADA7= "7DADA (°C)",
    Length_mean_sd = md("Length (mm)  \n ±sd"), #md required for line break to work
    LOE_mean_sd = md("LOE Temp(°C)  \n ±sd")
  ) %>% 
  tab_options(
    table.font.names = "Times New Roman"
  ) %>% 
  tab_style(
    style= list(
      cell_text(weight= "bold", font = "Times New Roman")
    ),
    locations = cells_column_labels(everything())
  )
gt_table

gtsave(gt_table, "output/LOE_Length_7DADM.png")
```
