---
title: "CKMRsim: A Population level analysis"
format: pdf
editor: visual
---

## CKMRsim: A Population level analysis

The last ckmr script analysed the whole data set at once. While this didn't initially appear to be a problem, we want to assume that each population is isolated from one another. As a result, we need to run ckmr again on each population.

1.  Load in libraries and data
2.  Create new data frames for each watershed/population
3.  Attempt to convert the different ckmr code chunks and steps into functions
4.  Apply functions (if successful) to each population

## 1. Prep library and data

```{r}
#| echo: false
library(tidyverse) 
library(MCMCpack) # for animal model
library(ggplot2) # plots
library(here) # easy data file access for Git
library(CKMRsim)
```

```{r}
# Phenotype and metadata
ctmax <- read.csv(here("Genomics/Data/Omy5_inversion_metadata_merge.csv"))

# SNP data -------------------------------------------------------------------
# Data received 7/2/26
snp <- read.csv(here("Genomics/Data/PaigeAdapCapFishSNPGenotypes.raw.csv"))
# extract first 7 characters into column with the same name as the ctmax dataframe
str(snp$Sample_ID)
snp$indiv.ID <- substr(snp$Sample_ID,1,7)
# remove original sample_ID column
snp <- snp %>% 
  dplyr::select(-Sample_ID)
snp <- snp %>% relocate(last_col())

### rename the columns ----------------------
# make the periods be _ for the next steps.
snp <- snp %>% 
  rename_with(~ gsub(".", "_",.x, fixed = TRUE))
# make note of the current order of loci in the data set
nc <- ncol(snp)
loci <- str_replace( # expects string, pattern, replacement
  names(snp)# returns the names of the columns
  [seq(2, nc, by = 2)], # creates a sequence of even numbers starting at 2 and ending at nc, stepping by 2
  "\\.\\.\\.[0-9]+$", "") # this cleans up column names. So I don't think this will impact my dataframe at all

#  reset the locus names
names(snp)[seq(2, nc, by = 2)] <- str_c(loci, "1", sep = ".") #put a .1 at the end of the first allele
names(snp)[seq(3, nc, by = 2)] <- str_c(loci, "2", sep = ".") # put a .2 at the end of the second allele
names(snp)
snp <- snp %>% 
  rename(indiv.ID = indiv_ID)


# MHap data----------------------------------------------------------------------
## Neutral
neutral_mhap <- read.csv(here("Genomics/Data/Neutral_genotypes_mhapcode_wide.csv"))
nc <- ncol(neutral_mhap)
loci <- str_replace( # expects string, pattern, replacement
  names(neutral_mhap)# returns the names of the columns
  [seq(2, nc, by = 2)], # creates a sequence of even numbers starting at 2 and ending at nc, stepping by 2
  "\\.\\.\\.[0-9]+$", "") # this cleans up column names. So I don't think this will impact my dataframe at all

#  reset the locus names
names(neutral_mhap)[seq(2, nc, by = 2)] <- str_c(loci, "1", sep = ".") #put a .1 at the end of the first allele
names(neutral_mhap)[seq(3, nc, by = 2)] <- str_c(loci, "2", sep = ".") # put a .2 at the end of the second allele
names(neutral_mhap)


## Adaptive
adap_mhap <- read.csv(here("Genomics/Data/Adaptive_genotypes_mhapcode_wide.csv"))

```

Merge the SNP, adaptive, and neutral mhap data frames.

```{r}
merged1 <- merge(adap_mhap,neutral_mhap, by = "indiv.ID") 
fullmerge <- merge(snp, merged1, by = "indiv.ID")
```

Save the above data frame for easy access in the repository:

```{r}
write.csv(fullmerge, "Data/merged_ALLgenetics_codewide.csv", row.names = FALSE)
```

Filter out individuals pre-identified in Genomics.PreliminaryAnalysis, the SexId columns. Eventually, we will only want to keep two (ish) of the Omy columns in the adap_mhap data frame.

```{r}
all <-  fullmerge %>% 
  filter(!indiv.ID %in% c("M145767","M145542", # matching indivs from scott creek and big creek
                       "M145524", "M145692")) %>% 
  dplyr::select(-c(SexID.1,SexID.2))
```

Merge location data from the metadata file with the above data file.

```{r}
populations <- ctmax %>% 
  rename(indiv.ID = NMFS_DNA_ID_use) %>% 
   filter(!indiv.ID %in% c("M145767","M145542", 
                       "M145524", "M145692")) %>% 
  dplyr::select(c(indiv.ID,WATERSHED)) %>% 
  left_join(., all, by ="indiv.ID")
```

## 2. Separate Data Frames for Each Population

```{r}
bigcreek <- populations %>% 
  filter(WATERSHED == "Big Creek")

scottcreek <- populations %>% 
  filter(WATERSHED == "Scott Creek")

carmelriver <- populations %>% 
  filter(WATERSHED == "Carmel River")

napariver <- populations %>% 
  filter(WATERSHED == "Napa River")

russianriver <- populations %>% 
  filter(WATERSHED == "Russian River")
```

## 3. CKMR functions

This is where I'm going to leave off for the day. Next, I will try to convert the CKMR code chunks into functions that can be easily applied to each location for ease of replication and consistency.
