---
title: "Genomics Prep"
format: html
editor: visual
---

## Genomics

There were 12 samples excluded from the SNP analysis due to sample quality.

1.  Trim up the dataframe and merge it with the phenotype data
2.  Which package to use for sibship analysis? Check Carlos' papers (maybe Anne's? Or maybe in Abadia-Cordoso paper)
3.  Quality check with some figures
4.  Build a matrix (ha. like it's simple)
5.  Quality check matrix a shit ton of times

```{r}
library(MCMCglmm)# for animal model
library(tidyverse) 
library(MCMCpack) # for animal model
library(ggplot2) # plots
library(here) # easy data file access for Git
library(adegenet)
```

```{r}
# Phenotype and metadata
ctmax <- read.csv(here("Genomics/Data/Omy5_inversion_metadata_merge.csv"))

# SNP data using 96 SNP panel from Abadia-Cordoso et al 2011
# Data received 7/2/26
snp <- read.csv(here("Genomics/Data/PaigeAdapCapFishSNPGenotypes.raw.csv"))
colnames(snp)
```

## Merge Dataframes

I need to merge these two data frames for Carlos. The only issue is Carlos says that 12 samples didn't run because they didn't pass quality check, but the same number of rows are present in each data frame. I'm not sure why that is- I will merge the data frames and ask Carlos.

First, I need to extract the first 7 characters from the Sample_ID column in the *snp* data frame to match the *ctmax* data frame.

```{r}
# extract first 7 characters into column with the same name as the ctmax dataframe
str(snp$Sample_ID)
snp$NMFS_DNA_ID_use <- substr(snp$Sample_ID,1,7)

# combine data frames by the NMFS_DNA_ID_use column
colnames(ctmax)
geno_pheno <- left_join(ctmax,snp, by ="NMFS_DNA_ID_use")

#write.csv(geno_pheno, "data/merged_PaigeAdapCapSNPs.csv", row.names = FALSE)
```

The new dataframe *geno_pheno* is the merged dataset we can now work with more easily.

Lastly, I exclude the sex chromosome and OmyAldA from the data frame, since that's adaptive and non-neutral.

```{r}
geno_pheno_neutral <- geno_pheno %>% 
  dplyr::select(-c(SexID,SexID.1))
```

## Developing a clear workflow

I think we will start with jumping into a genomic relatedness matrix.

Matrix type: VanRaden (<https://www.sciencedirect.com/science/article/pii/S0022030208709901?via%3Dihub>)

Notes from Abadia-Cordoso et al 2011:

-   SNP's are usually bi-allelic

-   Structure of the data:

    -   columns that end in ".1" is the pair. So, for example, 3 and 3 is homozygous, 1 and 3 is heterozygous

Let's give this matrix a go. I will start with the AGHmatrix package.

Workflow:

1.  Collapse alleles into SNP data: adegenet package should better be able to handle the data format
2.  Build GRM using AGHmatrix
3.  Do some quality checks- how?
4.  Invert the matrix - MCMCglmm (unless there's a clear way to do this in the AGHmatrix package)

## 1. Adegenet

### Install package

```{r}
# other useful packages, according to the tutorial I'm following online
library("ape")
library("pegas")
library("seqinr")
library("ggplot2")
```

#### *Adegenet notes*

There are two main classes of objects for storing genetic marker data, depending on the level at which the genetic information is considered:

-   genind for individual genotypes

-   genpop is used for alleles numbers counted by populations, where a population is any grouping of individuals

### HWE - Sanity Check

```{r}
expected_het <- gen_sum$Hexp
observed_het <- gen_sum$Hobs

# plot
plot(expected_het, observed_het,
     main = "Expected vs Observed Heterozygosity (Hexp)",
     xlab = "Expected Het",
     ylab = "Observed Het",
     pch = 20,
     col="blue")
abline(0,1,col = "red",lty=2)
```

This was a sanity check to see if what I had was in HWE.

**Question**: What is acceptable for HWE deviation? Is this concerning?

Check: HWE deviation in each watershed

## 2. CKMRsim: Look for Full- Sibs (From E. Anderson Git)

First, I will check for full-sibs in just the SNP data. Then I will add in the neutral mhaps and check for full-sibs again, then I will add in the adaptive mhaps too.

```{r}
library(CKMRsim)
library(tidyverse)
library(here)

snp1 <- snp %>% 
  dplyr::select(-c(SexID,SexID.1,Sample_ID)) %>% 
  dplyr::select(NMFS_DNA_ID_use, everything()) #make NMFS_DNA_ID_use the first column

```

This will be done using the dataframe called snp1, which has been appropriately cleaned.

### Estimating power for relationship inference

Estimating power for pairwise relationships involves MCMC simulations from the allele frequencies in the population. So first we have to estimate the allele freqs from our genotype data set. And we have to present those allele frequencies in a specific format that CKMRsim is expecting.

#### Allele frequency input format

Genome coordinates of markers are not required but the input format has a place to record the chromosome and position of each marker. If these are not known, we can use dummy values. The values should have no effect as long as there is no physical linkage.

Note that the following columns are expected:

-   Chrom: chromosome that the marker is on

-   Locus: name of markers/loci (no spaces!)

-   Pos: genome coordinates (numeric)

-   Allele: name of each allele. must be a character, so you might have to coerce numbers into characters

-   Freq: frequency of each allele in the populations. Should sum to 1 over each locus

```{r}
# make the periods be _ for the next steps.
snp1 <- snp1 %>% 
  rename_with(~ gsub(".", "_",.x, fixed = TRUE))
# make note of the current order of loci in the data set
nc <- ncol(snp1)
loci <- str_replace( # expects string, pattern, replacement
  names(snp1)# returns the names of the columns
  [seq(2, nc, by = 2)], # creates a sequence of even numbers starting at 2 and ending at nc, stepping by 2
  "\\.\\.\\.[0-9]+$", "") # this cleans up column names. So I don't think this will impact my dataframe at all

#  reset the locus names
names(snp1)[seq(2, nc, by = 2)] <- str_c(loci, "1", sep = ".") #put a .1 at the end of the first allele
names(snp1)[seq(3, nc, by = 2)] <- str_c(loci, "2", sep = ".") # put a .2 at the end of the second allele
names(snp1)

# then make some long format genotypes
long_genos <- snp1 %>% 
  gather(key = "loc", value = "Allele", -NMFS_DNA_ID_use) %>%
  separate(loc, into = c("Locus", "gene_copy"), sep = "\\.") %>%
  mutate(Allele = as.character(Allele)) %>%
  mutate(Allele = ifelse(Allele == "0", NA, Allele)) %>%
  rename(Indiv = NMFS_DNA_ID_use)
colnames(snp1)

str(long_genos)
unique(long_genos$Locus)
```

#### Calculate allele frequencies

```{r}
colnames(long_genos)
anyDuplicated(names(long_genos))
alle_freqs <- long_genos %>%
  count(Locus,Allele) %>%
  group_by(Locus) %>%
  mutate(Freq = n / sum(n),
         Chrom = "Unk",
         Pos = as.integer(factor(Locus, levels = loci))) %>%
  ungroup() %>%
  dplyr::select(Chrom, Pos, Locus, Allele, Freq) %>%
  arrange(Pos, desc(Freq)) %>%
  mutate(AlleIdx = NA, # leave this as NA
         LocIdx = NA) %>% # leave this as NA
  filter(!is.na(Allele)) # removing NA's is really important in this step!

# see what that looks like:
alle_freqs
```

Now we can make a dataframe that CKMRsim can use.

```{r}
afreqs_ready <- reindex_markers(alle_freqs)
```

#### Create CKMRsim object

```{r}
ckmr <- create_ckmr(
  D = afreqs_ready,
  kappa_matrix = kappas[c("PO", "FS", "HS", "U"), ],
  ge_mod_assumed = ge_model_TGIE, # function describes genotyping error model applied to model for computing likelihoods. TGIE is for integers
  ge_mod_true = ge_model_TGIE, # genotyping error model to simulate genotype data
  ge_mod_assumed_pars_list = list(epsilon = 0.005), # named parameters for assumed genotyping error model (can leave blank and use defaults)
  ge_mod_true_pars_list = list(epsilon = 0.005) # for true genotyping error
)

ckmr
```

This looks great so far! This is a bi-allelic data frame, so we feel good about loci having 2 alleles. Remember that this is an S3 object with class ckmr.

#### Simulate genotype pairs + calculate log-probabilities

```{r}
ex1_Qs <- simulate_Qij(ckmr, 
                       calc_relats = c("PO", "FS", "U"),
                       sim_relats = c("PO", "FS", "HS", "U") )
ex1_Qs
```

#### Compute log-likelihood ratios to view them

```{r}
FS_U_logls <- extract_logls(ex1_Qs,
                            numer = c(FS = 1),
                            denom = c(U = 1))
FS_U_logls
```

Visualize this in ggplot.

```{r}
ggplot(FS_U_logls,
            aes(x = logl_ratio, fill = true_relat)) +
  geom_density(alpha = 0.25)
```

From this, we see that there is little overlap in the FS/U log likelihood ratios.

```{r}
ggplot(FS_U_logls %>% filter(true_relat %in% c("FS", "U")),
            aes(x = logl_ratio, fill = true_relat)) +
  geom_density(alpha = 0.25)
```

#### Estimating false negative and positive rates

```{r}
ex1_FS_is <- mc_sample_simple(ex1_Qs, 
                 nu = "FS",
                 de = "U")

ex1_FS_is

# question: should I check FS against HS?
```

The tutorial for CKMRsim recommends requiring that the FPR (false positive rate) be about 10-100 times smaller than the reciprocal of the number of comparisons. In this case, thats

```{r}
0.1*((494*493)/2) ^-1
```

```{r}
ex1_PO_is_8_12 <- mc_sample_simple(ex1_Qs, 
                 nu = "FS",
                 de = "U", 
                 lambda_stars = seq(12, 15, by = .1))
ex1_PO_is_8_12
```

We should set logl to at least 15 to accomplish what the authors recommend for FPR.

#### Screen for duplicate samples

```{r}
matchers <- find_close_matching_genotypes(LG = long_genos,
                                          CK = ckmr,
                                          max_mismatch = 6)

matchers

pairs_matchers <- matchers %>%
  left_join(
    ctmax %>%
      dplyr::select(NMFS_DNA_ID_use, WATERSHED,LENGTH,COLLECTION_DATE,SAMPLE_ID) %>%
      rename(indiv_1 = NMFS_DNA_ID_use,
             Watershed_1 = WATERSHED,
             Length_1 =LENGTH,
             Collection_Date_1 = COLLECTION_DATE,
             Sample_ID_1 = SAMPLE_ID),
    by = "indiv_1"
  ) %>%
  left_join(
    ctmax %>%
      dplyr::select(NMFS_DNA_ID_use, WATERSHED,LENGTH,COLLECTION_DATE,SAMPLE_ID) %>%
      rename(indiv_2 = NMFS_DNA_ID_use,
             Watershed_2 = WATERSHED,
             Length_2 =LENGTH,
             Collection_Date_2 = COLLECTION_DATE,
             Sample_ID_2 = SAMPLE_ID),
    by = "indiv_2"
  )
pairs_matchers
```

I found three pairs of matches. It's strange that a Big Creek and Scott Creek individual matched- I will remove both individuals from the data set.

I will keep the first (indiv_1) from the other two pairs (M145494 and M145595)

```{r}
write.csv(pairs_matchers, "Data/snps_matchedindiv_ckmr.csv", row.names = FALSE)
```

#### Filter out matchers

```{r}
long_genos_filtered <- long_genos %>% 
  filter(!Indiv %in% c("M145767","M145542", # matching indivs from scott creek and big creek
                       "M145524", "M145692"))
```

#### Compute logl ratios for pairwise comparisons

```{r}
fs_pairwise_logls <- pairwise_kin_logl_ratios(D1 = long_genos_filtered,
                                              D2 = long_genos_filtered,
                                              CK = ckmr,
                                              numer = "FS",
                                              denom = "U", 
                                              num_cores = 1)

# note, num_cores is set to 1 because more than that breaks
# it on CRAN checks.  But, leave it blank (i.e., don't use the
# num_core option) to, by default, use all
# your cores in parallel on a non-windows machine.

# I used log 12 bc that's what Kerry used

fs_pairwise_logls12 <- fs_pairwise_logls %>%
  filter(logl_ratio > 12) %>%
  arrange(desc(logl_ratio))
```

```{r}
write.csv(fs_pairwise_logls12,"Data/snps_fs_pairwise_logls12.csv")
```



## 3. Visualize CKMRsim output

### Merge sibling data frame with metadata

```{r}
colnames(fs_pairwise_logls)
colnames(ctmax)

# chatGPT helped with this block
pairs_FS12 <- fs_pairwise_logls12 %>%
  left_join(
    ctmax %>%
      dplyr::select(NMFS_DNA_ID_use, WATERSHED,LENGTH,COLLECTION_DATE,SAMPLE_ID) %>%
      rename(D1_indiv = NMFS_DNA_ID_use,
             Watershed_1 = WATERSHED,
             Length_1 =LENGTH,
             Collection_Date_1 = COLLECTION_DATE,
             Sample_ID_1 = SAMPLE_ID),
    by = "D1_indiv"
  ) %>%
  left_join(
    ctmax %>%
      dplyr::select(NMFS_DNA_ID_use, WATERSHED,LENGTH,COLLECTION_DATE,SAMPLE_ID) %>%
      rename(D2_indiv = NMFS_DNA_ID_use,
             Watershed_2 = WATERSHED,
             Length_2 =LENGTH,
             Collection_Date_2 = COLLECTION_DATE,
             Sample_ID_2 = SAMPLE_ID),
    by = "D2_indiv"
  )

# add a column to check if watersheds are the same
pairs_FS12 <- pairs_FS12 %>%
  mutate(Watersheds_same = Watershed_1 == Watershed_2)
```

#### Heatmap of relatedness

Not entirely sure how to do this yet- maybe geom_tile from ggplot2 or pheatmap

To keep progressing, I'm going to skip this for now. I'm not entirely sure what should go on the diagonal/ 1:1 line...

### 4. Add in Neutral mhaps to CKMRsim

Ok so we have now assessed full-sibling pairs using CKMRsim on just the SNP data. However, we are eventually going to use all the mhap data for the matrix, so this next step combines neutral mhap data with the snp data, and follows the same steps as above. Remember that mhaps are multi-allelic. I don't think that should cause any problems at this stage, but it will have to be dealt with before we build the matrix.

#### Load in neutral mhap data frame and prep

```{r}
mhap1 <- read.csv(here("Genomics/Data/Neutral_genotypes_mhapcode_wide.csv"))
```

Remove the same individuals removed above:

```{r}
mhap1 <- mhap1 %>% 
  filter(!indiv.ID %in% c("M145767","M145542", # matching indivs from scott creek and big creek
                       "M145524", "M145692"))
```

Calculate allele frequencies

```{r}
nc <- ncol(mhap1)
loci <- str_replace( # expects string, pattern, replacement
  names(mhap1)# returns the names of the columns
  [seq(2, nc, by = 2)], # creates a sequence of even numbers starting at 2 and ending at nc, stepping by 2
  "\\.\\.\\.[0-9]+$", "") # this cleans up column names. So I don't think this will impact my dataframe at all

#  reset the locus names
names(mhap1)[seq(2, nc, by = 2)] <- str_c(loci, "1", sep = ".") #put a .1 at the end of the first allele
names(mhap1)[seq(3, nc, by = 2)] <- str_c(loci, "2", sep = ".") # put a .2 at the end of the second allele
names(mhap1)

# then make some long format genotypes
mhaplong_genos <- mhap1 %>% 
  gather(key = "loc", value = "Allele", -indiv.ID) %>%
  separate(loc, into = c("Locus", "gene_copy"), sep = "\\.") %>%
  mutate(Allele = as.character(Allele)) %>%
  mutate(Allele = ifelse(Allele == "0", NA, Allele)) %>%
  rename(Indiv = indiv.ID)
colnames(mhaplong_genos)

str(mhaplong_genos)
unique(mhaplong_genos$Locus)
```

Now I think we can just go ahead and combine data frames.

```{r}
mhap_snp <- bind_rows(long_genos_filtered, mhaplong_genos)
```

#### Calculate allele freqs

```{r}
alle_freqs2 <- mhap_snp %>%
  count(Locus,Allele) %>%
  group_by(Locus) %>%
  mutate(Freq = n / sum(n),
         Chrom = "Unk",
         Pos = as.integer(factor(Locus, levels = loci))) %>%
  ungroup() %>%
  dplyr::select(Chrom, Pos, Locus, Allele, Freq) %>%
  arrange(Pos, desc(Freq)) %>%
  mutate(AlleIdx = NA, # leave this as NA
         LocIdx = NA) %>% # leave this as NA
  filter(!is.na(Allele)) # removing NA's is really important in this step!

# see what that looks like:
alle_freqs2
```

Make a CKRMsim object

```{r}
afreqs_ready2 <- reindex_markers(alle_freqs2)
```

#### Create CKRMsim object

```{r}
ckmr2 <- create_ckmr(
  D = afreqs_ready2,
  kappa_matrix = kappas[c("PO", "FS", "HS", "U"), ],
  ge_mod_assumed = ge_model_TGIE, # function describes genotyping error model applied to model for computing likelihoods. TGIE is for integers
  ge_mod_true = ge_model_TGIE, # genotyping error model to simulate genotype data
  ge_mod_assumed_pars_list = list(epsilon = 0.005), # named parameters for assumed genotyping error model (can leave blank and use defaults)
  ge_mod_true_pars_list = list(epsilon = 0.005) # for true genotyping error
)

ckmr2
```

#### Simulate genotype pairs + calculate log-probabilities

```{r}
ex1_Qs2 <- simulate_Qij(ckmr2, 
                       calc_relats = c("PO", "FS", "U"),
                       sim_relats = c("PO", "FS", "HS", "U") )
ex1_Qs2

mhap_snp_PO_is_8_12 <- mc_sample_simple(ex1_Qs2, 
                 nu = "FS",
                 de = "U")

ex1_PO_is_8_12_2 <- mc_sample_simple(ex1_Qs2, 
                 nu = "FS",
                 de = "U", 
                 lambda_stars = seq(12, 15, by = .1))

ex1_PO_is_8_12_2
```

#### Compute log-likelihood ratios to view them

```{r}
FS_U_logls2 <- extract_logls(ex1_Qs2,
                            numer = c(FS = 1),
                            denom = c(U = 1))
FS_U_logls2
```

Visualize this in ggplot.

```{r}
ggplot(FS_U_logls2,
            aes(x = logl_ratio, fill = true_relat)) +
  geom_density(alpha = 0.25)
```

```{r}
ggplot(FS_U_logls2 %>% filter(true_relat %in% c("FS", "U")),
            aes(x = logl_ratio, fill = true_relat)) +
  geom_density(alpha = 0.25)
```

*Consider checking for matchers again with the two dataframes combined.*

```{r}
fs_pairwise_logls2 <- pairwise_kin_logl_ratios(D1 = mhap_snp,
                                              D2 = mhap_snp,
                                              CK = ckmr2,
                                              numer = "FS",
                                              denom = "U", 
                                              num_cores = 1)

fs_pairwise_logls12_mhap_snp <- fs_pairwise_logls2 %>%
  filter(logl_ratio > 15) %>%
  arrange(desc(logl_ratio))

fs_pairwise_logls_mhap_snp <- fs_pairwise_logls2 %>%
 # filter(logl_ratio > 12) %>%
  arrange(desc(logl_ratio))

fs_pairwise_logls12_mhap_snp

PO_U_logls <- extract_logls(ex1_Qs2,
                            numer = c(FS = 1),
                            denom = c(U = 1))

ex1_PO_is_8_12 %>%
  filter(FPR > 1.275727e-08) %>%
  arrange(FPR) %>%
  slice(1)
```

I think something strange happened here- I found 2.5X full sibs that I found in the snp analysis alone. I would have expected less, I think.

PICKING UP Why are there more full-sibs found in the new dataframe compared to the old? Maybe 12 is wrong?


## 5. ALL Data in CKMRsim 

I've now done the CKMRsim analyses using SNP and SNP + neutral mhaps. Lastly, I'll add in the adaptive mhaps.

There are 5 mhap markers associated with Omy05 and 2-3 SNPs. We will want to omit eventually all but 2-3 of these to avoid pseudo replication.

```{r}
# Read in adaptive panel csv
adap_mhap <- read.csv(here("Genomics/Data/Adaptive_genotypes_mhapcode_wide.csv"))
length(unique(adap_mhap))

# Just noticing that a lot of this data frame is 1 and 2 - I guess this kind of makes sense? No major changes in the major adaptive regions
```

Clean the csv by filtering out the same individuals as above and merge it with the other data frame.

```{r}
adap_mhap_filtered <- adap_mhap %>% 
  rename(Indiv = indiv.ID) %>% # rename indiv column to match
  filter(!Indiv %in% c("M145767","M145542", # matching indivs from scott creek and big creek
                       "M145524", "M145692"))
```


### Prepare the data frame
```{r}
nc <- ncol(adap_mhap_filtered)
loci <- str_replace( # expects string, pattern, replacement
  names(adap_mhap_filtered)# returns the names of the columns
  [seq(2, nc, by = 2)], # creates a sequence of even numbers starting at 2 and ending at nc, stepping by 2
  "\\.\\.\\.[0-9]+$", "") # this cleans up column names. So I don't think this will impact my dataframe at all


# long format
long_adap_mhap <- adap_mhap_filtered %>% 
  gather(
    key = "loc", 
    value = "Allele",
    -Indiv
  ) %>%
  separate(
    loc,
    into = c("Locus", "gene_copy"),
    sep = "\\.(?=[^.]+$)" # chat wrote this section
  ) %>%
  mutate(
    Allele = as.character(Allele),
    Allele = ifelse(Allele == "0", NA, Allele)
  )

# this seems to have worked, but could probably use a sanity check (chat helped trouble shoot it)
```

### Combine long data frames and Calculate allele frequencies

```{r}
all_data_long <- bind_rows(mhap_snp, long_adap_mhap)
```

#### Calculate allele freqs

```{r}
alle_freqs_all<- all_data_long %>%
  count(Locus,Allele) %>%
  group_by(Locus) %>%
  mutate(Freq = n / sum(n),
         Chrom = "Unk",
         Pos = as.integer(factor(Locus, levels = loci))) %>%
  ungroup() %>%
  dplyr::select(Chrom, Pos, Locus, Allele, Freq) %>%
  arrange(Pos, desc(Freq)) %>%
  mutate(AlleIdx = NA, # leave this as NA
         LocIdx = NA) %>% # leave this as NA
  filter(!is.na(Allele)) # removing NA's is really important in this step!

# see what that looks like:
alle_freqs_all
```

Make a CKRMsim object

```{r}
afreqs_ready_all <- reindex_markers(alle_freqs_all)
```

#### Create CKRMsim object

```{r}
ckmr_all <- create_ckmr(
  D = afreqs_ready_all,
  kappa_matrix = kappas[c("PO", "FS", "HS", "U"), ],
  ge_mod_assumed = ge_model_TGIE, # function describes genotyping error model applied to model for computing likelihoods. TGIE is for integers
  ge_mod_true = ge_model_TGIE, # genotyping error model to simulate genotype data
  ge_mod_assumed_pars_list = list(epsilon = 0.005), # named parameters for assumed genotyping error model (can leave blank and use defaults)
  ge_mod_true_pars_list = list(epsilon = 0.005) # for true genotyping error
)

ckmr_all
```


#### Simulate genotype pairs + calculate log-probabilities

```{r}
sim_all <- simulate_Qij(ckmr_all, 
                       calc_relats = c("PO", "FS", "U"),
                       sim_relats = c("PO", "FS", "HS", "U") )
sim_all

ALL_PO_is_8_12 <- mc_sample_simple(sim_all,
                 nu = "FS",
                 de = "U")
ALL_PO_is_8_12

sim_all_FPR <- mc_sample_simple(sim_all, 
                 nu = "FS",
                 de = "U", 
                 lambda_stars = seq(12, 25, by = .1))
```
#### Compute log-likelihood ratios to view them

```{r}
FS_U_logls_ALL <- extract_logls(sim_all,
                            numer = c(FS = 1),
                            denom = c(U = 1))
FS_U_logls_ALL
```

Visualize this in ggplot.

```{r}
ggplot(FS_U_logls_ALL,
            aes(x = logl_ratio, fill = true_relat)) +
  geom_density(alpha = 0.25)
```
### Check for number of full sib pairs
```{r}
fs_pairwise_logls_ALL <- pairwise_kin_logl_ratios(D1 = all_data_long,
                                              D2 = all_data_long,
                                              CK = ckmr_all,
                                              numer = "FS",
                                              denom = "U", 
                                              num_cores = 1)

FS_FPR_is <- mc_sample_simple(sim_all, 
                 nu = "FS",
                 de = "U")

FS_FPR_is

FS_FPR_is %>%
  filter(FPR > 8.2e-07) %>%
  arrange(FPR) %>%
  slice(1)

fs_pairwise_logls_ALL <- fs_pairwise_logls_ALL %>%
  filter(logl_ratio > 20) %>%
  arrange(desc(logl_ratio))
```

Use a lambda star of 18 (uncertain here). There are only 5 rows, and all rows comply with using a lambda of 18. I think this indicates that we have really good power. 


```{r}
.1*((494*493)/2)^-1
```

# 6. Visualization of Data

I am going to make a pairwise plot that compares individuals to one another, but sorts by location and year, using logLR >18.

First, I need to add in metadata to each individual. 

```{r}
# rename indiv columns
pairwise_sibs <- fs_pairwise_logls_ALL %>% 
  rename(indiv_1 = D1_indiv,
         indiv_2 = D2_indiv)

# clean ctmax dataframe
ctmax_clean <- ctmax %>% 
  mutate(COLLECTION_DATE = mdy(COLLECTION_DATE)) %>% 
  mutate(YEAR = year(COLLECTION_DATE))

# two merges:
pairwise_sibsmerge1 <- pairwise_sibs %>% 
  left_join(ctmax_clean %>% 
              dplyr::select(NMFS_DNA_ID, WATERSHED, YEAR) %>%
              rename(watershed_1 = WATERSHED,
                     year_1 = YEAR,
                     indiv_1 = NMFS_DNA_ID),
            by = "indiv_1")

#merge 2
pairwise_sibsmerged <- pairwise_sibsmerge1 %>% 
  left_join(ctmax_clean %>% 
              dplyr::select(NMFS_DNA_ID, WATERSHED, YEAR) %>%
              rename(watershed_2 = WATERSHED,
                     year_2 = YEAR,
                     indiv_2 = NMFS_DNA_ID),
            by = "indiv_2")
```


I think we should only retain individuals that are FS. So where logLR >18.
```{r}
filtered_sibs <- pairwise_sibsmerged %>%
  filter(logl_ratio > 12) %>%
  arrange(desc(logl_ratio))

# pairwise relatedness column
pairwise_sibsmerged <- pairwise_sibsmerged %>%
  mutate(
    comparison = case_when(
      watershed_1 == watershed_2 ~ "Within location",
      TRUE ~ "Between locations"
    )
  )


# Get all individuals and order them by location and year
individuals <- ctmax_clean %>% 
  arrange(WATERSHED, YEAR, NMFS_DNA_ID)

# this is where I left off

matrix_data <- individuals %>%
  select(ID1 = ID, location1 = location, year1 = study_year, order1 = order) %>%
  inner_join(
    individuals %>%
      select(ID2 = ID, location2 = location, year2 = study_year, order2 = order),
    by = character()
  ) %>%
  filter(location1 == location2)

ggplot(matrix_data, aes(x = order1, y = order2, fill = relatedness)) +
  geom_tile() +
  geom_abline(
    slope = 1,
    intercept = 0,
    linewidth = 0.5
  ) +
  scale_fill_viridis_c(
    name = "Genetic\nrelatedness",
    na.value = "white"
  ) +
  scale_x_continuous(
    breaks = individuals$order,
    labels = individuals$ID,
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    breaks = individuals$order,
    labels = individuals$ID,
    expand = c(0, 0)
  ) +
  coord_equal() +
  theme_classic() +
  theme(
    axis.text.x = element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.5
    ),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )
```


```{r}

```






