---
title: "Genomics Prep"
format: html
editor: visual
---

## Genomics

There were 12 samples excluded from the SNP analysis due to sample quality.

1.  Trim up the dataframe and merge it with the phenotype data
2.  Which package to use for sibship analysis? Check Carlos' papers (maybe Anne's? Or maybe in Abadia-Cordoso paper)
3.  Quality check with some figures
4.  Build a matrix (ha. like it's simple)
5.  Quality check matrix a shit ton of times

```{r}
library(MCMCglmm)# for animal model
library(tidyverse) 
library(MCMCpack) # for animal model
library(ggplot2) # plots
library(here) # easy data file access for Git
library(adegenet)
```

```{r}
# Phenotype and metadata
ctmax <- read.csv(here("Genomics/Data/Omy5_inversion_metadata_merge.csv"))

# SNP data using 96 SNP panel from Abadia-Cordoso et al 2011
# Data received 7/2/26
snp <- read.csv(here("Genomics/Data/PaigeAdapCapFishSNPGenotypes.raw.csv"))
colnames(snp)
```

(

## Merge Dataframes

I need to merge these two data frames for Carlos. The only issue is Carlos says that 12 samples didn't run because they didn't pass quality check, but the same number of rows are present in each data frame. I'm not sure why that is- I will merge the data frames and ask Carlos.

First, I need to extract the first 7 characters from the Sample_ID column in the *snp* data frame to match the *ctmax* data frame.

```{r}
# extract first 7 characters into column with the same name as the ctmax dataframe
str(snp$Sample_ID)
snp$NMFS_DNA_ID_use <- substr(snp$Sample_ID,1,7)

# combine data frames by the NMFS_DNA_ID_use column
colnames(ctmax)
geno_pheno <- left_join(ctmax,snp, by ="NMFS_DNA_ID_use")

#write.csv(geno_pheno, "data/merged_PaigeAdapCapSNPs.csv", row.names = FALSE)
```

The new dataframe *geno_pheno* is the merged dataset we can now work with more easily.

Lastly, I exclude the sex chromosome and OmyAldA from the data frame, since that's adaptive and non-neutral.

```{r}
geno_pheno_neutral <- geno_pheno %>% 
  dplyr::select(-c(SexID,SexID.1))
```

## Developing a clear workflow

I think we will start with jumping into a genomic relatedness matrix.

Matrix type: VanRaden (<https://www.sciencedirect.com/science/article/pii/S0022030208709901?via%3Dihub>)

Notes from Abadia-Cordoso et al 2011:

-   SNP's are usually bi-allelic

-   Structure of the data:

    -   columns that end in ".1" is the pair. So, for example, 3 and 3 is homozygous, 1 and 3 is heterozygous

Let's give this matrix a go. I will start with the AGHmatrix package.

Workflow:

1.  Collapse alleles into SNP data: adegenet package should better be able to handle the data format
2.  Build GRM using AGHmatrix
3.  Do some quality checks- how?
4.  Invert the matrix - MCMCglmm (unless there's a clear way to do this in the AGHmatrix package)

## 1. Adegenet

### Install package

```{r}
# other useful packages, according to the tutorial I'm following online
library("ape")
library("pegas")
library("seqinr")
library("ggplot2")
```

#### *Adegenet notes*

There are two main classes of objects for storing genetic marker data, depending on the level at which the genetic information is considered:

-   genind for individual genotypes

-   genpop is used for alleles numbers counted by populations, where a population is any grouping of individuals

### Manually designating 0,1,2 snps

I'm pretty sure all this is wrong. I think adegenet can handle this, the columns just need to be renamed.

Don't forget to get rid of sex id column!!! And check for other adaptive alleles (ask Ellen?)

Now create a genind object for adegenet

```{r}
# replace all decimal points with a "_", otherwise adegenet doesn't know how to read it
snp_neutral <- snp %>% 
  dplyr::select(-c(SexID,SexID.1))

# rename columns so that genind can better read them
colnames(snp_neutral)<- gsub(".", "_", colnames(snp_neutral), fixed = TRUE)
colnames(snp_neutral)<- gsub("_1", "_2", colnames(snp_neutral), fixed = TRUE)
colnames(snp_neutral) <- ifelse(
  grepl("")
)

gen <- df2genind(
  geno_df,
  sep = "/",
  ind.names = ids,
  ploidy = 2,
  type = "codom",
  strata = 5
)
?df2genind

gen
gen_sum <- summary(gen)
```

### HWE - Sanity Check

```{r}
expected_het <- gen_sum$Hexp
observed_het <- gen_sum$Hobs

# plot
plot(expected_het, observed_het,
     main = "Expected vs Observed Heterozygosity (Hexp)",
     xlab = "Expected Het",
     ylab = "Observed Het",
     pch = 20,
     col="blue")
abline(0,1,col = "red",lty=2)
```

This was a sanity check to see if what I had was in HWE.

**Question**: What is acceptable for HWE deviation? Is this concerning?

Check: HWE deviation in each watershed

### CKMRsim: Look for Full- Sibs

I don't think we need to exclude full-sibs from the animal model, but it will still be helpful as a sanity check.

### DAPC Plot - Sanity Check

The genind object made above didn't work properly. For example, we should have 5 groupings. This is important for the DAPC.

From ChatGPT:

```{r}
# first, decide number of PC's to run
xval <- xvalDapc(
    gen,
    pop(gen),
    training.set = 0.9,
    result = "groupMean",
    center = TRUE,
    scale = FALSE,
    n.pca.max = 100
)
table(pop(gen))
```

Left off: I now need to code this object so that AGHmatrix can read it. So I need every combination to be either a 0,1, or 2, depending on homo vs heterozygosity.
