---
title: "Compressing multi allelic microhaplotypes"
format: html
editor: visual
---

## Compressing multi-allelic microhaplotypes to bi-allelic format

Building a genomic relatedness matrix (GRM) requires bi-allelic data. This is not compatible with our microhaplotype data, which sometimes has as many as 8 forms of one allele. Instead, we can compress this using a greedy algorithm. The greedy algorithm is a problem-solving algorithm that picks the most optimal choice at each step in the process. It does not consider previous choices- each choice is unique.

E. Anderson wrote the following code for A. Beulke for her PhD research. He gave her permission to share that file with me. I have adapted it a bit with additional comments to help my understanding and any replication of this process.

A concept he references frequently is minor allele frequencies (MAF's). We are shooting for MAF's of 0.5- this would show that the algorithm selected the most likely alleles at each step, thus working properly.

```{r}
library(tidyverse)
```

This is E. Anderson's code. First, he writes a function that acts any non-biallelic (SNP, or simple mhaps) into the algorithm.

```{r}
partition_greedy <- function(x) {
  
  if(length(x) == 2) {  # if it is already biallelic, just leave it
    return(
      tibble( # create a new data frame ewith tibble with the following col names
        AlleleOrig = names(x), # these are new column names we're creating for alleles that are already biallelic
        LumpedAllele = names(x),
        FreqOrig = unname(x) # strips out row and column labels
      )
    )
  } 
  # Sort descending to place large items first
  x_sorted <- sort(x, decreasing = TRUE)
  
  group1 <- c()
  group2 <- c()
  sum1 <- 0
  sum2 <- 0
  
  for (i in 1:length(x_sorted)) {
    val = x_sorted[i]
    if (sum1 <= sum2) {
      group1 <- c(group1, val)
      sum1 <- sum1 + unname(val)
    } else {
      group2 <- c(group2, val)
      sum2 <- sum2 + unname(val)
    }
  }
  
  # we use V and W to refer to the different lumped groups
  tibble(
    AlleleOrig = c(names(group1), names(group2)),
    LumpedAllele = c(rep("V", length(group1)), rep("W", length(group2))),
    FreqOrig = unname(c(group1, group2))
  )
}

```

Now I'll read in my first .rds file. This file is just for Big Creek.

```{r}
bigcreek_freq = read_rds("Data/bigcreek_allelefreqs.rds")
str(bigcreek_freq$Allele)

lumpy <- bigcreek_freq %>%
  group_by(Chrom, Pos, Locus) %>%
  nest() %>%
  mutate(
    lumps = map(.x = data, .f = function(x) {
      v <- x$Freq
      names(v) <- x$Allele
      partition_greedy(v)
    })
  ) %>%
  unnest(cols = c(data, lumps)) %>%
  ungroup()

lumpy
```

I don't have AlleIdx or LocIdx, which Anne had. And I wonder if that's because my input data frame doesn't have the original genotypes in them. This wasn't a problem for the CKMRsim output, so I'll push forward, knowing this could be an issue later.

Now we look at how many alleles we end up with:

```{r}
new_freqs <- lumpy %>%
  group_by(Chrom, Pos, Locus, LumpedAllele) %>%
  summarise(
    LumpedFreq = sum(FreqOrig),
    orig_num_alle = n()
  )

new_freqs
```

At this point, my lumpedallele column is still using numbers, whereas his example has V and W. It still seemed to work, but this may be worth changing.

```{r}
MAFs <- new_freqs %>%
  filter(LumpedFreq == min(LumpedFreq)) %>%
  rename(MAF = LumpedFreq) %>%
  mutate(
    type = case_when(
      str_detect(Locus, "(hap1_code|haplotype)$") ~ "mhap",
      TRUE ~ "SNP"
    )
  )

ggplot(MAFs, aes(x = MAF, fill = type)) +
  geom_histogram(binwidth = 0.01) +
  facet_wrap(~ type, ncol = 1)
```

This didn't turn out nearly as pretty as E. Anderson's did on Anne's data. I'm not sure if something went wrong. I will proceed as if it's ok, but go over this in depth with Kerry once she returns from leave.

```{r}
joiner = lumpy %>% 
  dplyr::select(Chrom, Pos, Locus, AlleleOrig, LumpedAllele)
```

This data frame can then be joined with the long_genos data frame created in the the CKMR.Populations file. E. Anderson recommends using by = join_by(Chrom, Pos, Locus, Allele == AlleleOrig)

## Read in long genos from Big Creek

```{r}
lg <- read_rds("Data/bigcreek_longgenos.rds")
```

Join by locus? I don't fully understand how we're not losing data here.

```{r}
bigcreek_lumpy = left_join(
  lg, joiner, 
  by =  join_by(Locus, Allele == AlleleOrig)) %>% 
  dplyr::select(-Allele) %>%  # need to delete old allele column
  rename(Allele = LumpedAllele) # replace with new allele column
```

Now I'm going to go back through ckmrsim and see what happens.

Calculate allele frequencies:

```{r}
bigcreekalle_freqs <- bigcreek_lumpy %>%
  dplyr::select(-Allele) %>%  # need to delete old allele column
  rename(Allele = LumpedAllele) %>% # replace with new allele column
  count(Locus, Allele) %>%
  group_by(Locus) %>%
  mutate(Freq = n / sum(n),
         Chrom = "Unk",
         Pos = as.integer(factor(Locus, levels = loci))) %>%
  ungroup() %>%
  dplyr::select(Chrom, Pos, Locus, Allele, Freq) %>%
  arrange(Pos, desc(Freq)) %>%
  mutate(AlleIdx = NA,
         LocIdx = NA) %>%
  filter(!is.na(Allele))
```

```{r}
afreqs_ready <- reindex_markers(bigcreekalle_freqs)
```

Create a ckmr object

```{r}
bigcreek_ckmr_lumpy <- create_ckmr(
  D = afreqs_ready,
  kappa_matrix = kappas[c("PO","FS","U","HS"), ],
  ge_mod_assumed = ge_model_TGIE,
  ge_mod_true = ge_model_TGIE,
  ge_mod_assumed_pars_list = list(epsilon = 0.005),
  ge_mod_true_pars_list = list(epsilon = 0.005)
)
bigcreek_ckmr_lumpy
```

Nice, so the greedy algorithm worked! The only concern is it says that it has between 2 and 2 alleles, and I think it should say between 1 and 2 alleles...

```{r}
bigcreek_Qs <- simulate_Qij(bigcreek_ckmr_lumpy, 
                       calc_relats = c("FS", "U","PO"),
                       sim_relats = c("FS",  "U","HS","PO") )
bigcreek_Qs
```

**Compute log-likelihood ratios to view them**

```{r}
bigcreek_FS_U_logls <- extract_logls(bigcreek_Qs,
                            numer = c(FS = 1),
                            denom = c(U = 1))

bigcreek_FS_U_logls
```

**Visualize**

```{r}
ggplot(bigcreek_FS_U_logls,
            aes(x = logl_ratio, fill = true_relat)) +
  geom_density(alpha = 0.25)
```

Check to make sure there are truly no identical individuals (there shouldn't be):

```{r}

bigcreek_lumpy_long_genos = bigcreek_lumpy %>% 
  dplyr::select(-c(Pos,Chrom))
matchers <- find_close_matching_genotypes(LG = bigcreek_lumpy_long_genos,
                                          CK = bigcreek_ckmr_lumpy,
                                          max_mismatch = 6)
matchers
```

Alright, so it worked!! Great! Let's see how many full sibs we have.

First, set loglr:

```{r}
bigcreek_FS_is <- mc_sample_simple(bigcreek_Qs, 
                 nu = "FS",
                 de = "U")

bigcreek_FS_is

#find where FNR > 0.01
bigcreek_FS_is %>%
  filter(FNR == 0.01) %>%
  arrange(FNR) %>%
  slice(1)

```

So logLR is 10. This makes some sense- we've collapsed some of the mhaps that gave us differentiation, so we lose some power.

```{r}
bigcreek_lumpy_pairwise_logls <- pairwise_kin_logl_ratios(
                                              D1 = bigcreek_lumpy_long_genos,
                                              D2 = bigcreek_lumpy_long_genos,
                                              CK = bigcreek_ckmr_lumpy,
                                              numer = "FS",
                                              denom = "U", 
                                              num_cores = 1)
```

Pairs with logLR greater than 10:

```{r}
bigcreek_lumpy_pairwise_logls %>%
  filter(logl_ratio > 10) %>%
  arrange(desc(logl_ratio))
```

We find 262 pairs. What about trios?

The next steps will be to build a sibling network and pull out groups. The goal should be to develop a workflow for each population that can be replicated for all other populations. During this network analysis, it will be important to consider year, as well.
