---
title: "Multi to biallelic"
format: html
editor: visual
---

## Prepping SNP/mhap data for GRM

There are three key things that need to happen for the SNP and microhaplotype data to be prepped for constructing the GRM. The key thing to remember is that both data types need to be in 0, 1, and 2 allele dosage format, where dosage counts the number alleles of each type.

1.  Convert SNP data to allele dosage. I will make a function to do this, which can be reused.
2.  Count unique microhaplotypes for each mhap
3.  Convert mhaps into allele dosage, where each unique mhap is its own column.

Once both data sets are in that format, they are combined into one data set.

## Structured Plan

1.  import SNP and mhap csv's, separately.
2.  Format column names properly. I believe they should end in \_#
3.  Convert SNPs into dosages
4.  Find each unique mhap at each locus
5.  Expand into dosage columns

Since there may be a lot of columns per mhap locus, it might be beneficial to weight the alleles.

```{r}
library(here)
library(tidyverse)
```

### 1. Import and clean .csv files

```{r}
# SNP
snp_raw <- read.csv(here("Genomics/Data/PaigeAdapCapFishSNPGenotypes.raw.csv"))

# Neutral mhap
neutral_mhap <- read.csv(here("Genomics/Data/Neutral_genotypes_mhapcode_wide.csv"))

# Adaptive mhap
adap_mhap <- read.csv(here("Genomics/Data/Adaptive_haplotypes_2col_wide.csv"))

# merged adaptive and neutral mhap data frames
all_mhap <- merge(neutral_mhap, adap_mhap, by = "indiv.ID")
```

Get rid of the sex ID column in the SNP data frame.

```{r}
snp <- snp_raw %>% 
  dplyr::select(-c(SexID,SexID.1))
```

### 2. Format columns

Reformatting columns for SNPs:

```{r}
snp <- snp %>% 
  rename_with(~ gsub(".", "_",.x, fixed = TRUE))
# make note of the current order of loci in the data set
nc <- ncol(snp)
loci <- str_replace( # expects string, pattern, replacement
  names(snp)# returns the names of the columns
  [seq(2, nc, by = 2)], # creates a sequence of even numbers starting at 2 and ending at nc, stepping by 2
  "\\.\\.\\.[0-9]+$", "") # this cleans up column names. So I don't think this will impact my dataframe at all

#  reset the locus names
names(snp)[seq(2, nc, by = 2)] <- str_c(loci, "1", sep = ".") #put a .1 at the end of the first allele
names(snp)[seq(3, nc, by = 2)] <- str_c(loci, "2", sep = ".") # put a .2 at the end of the second allele
names(snp)

# rename Sample_ID column to match mhaps. so that's indiv.ID
snp <- snp %>% 
  rename(indiv.ID = Sample_ID)
```

Reformatting columns for mhaps:

I left off here. The column names need to match the format of the SNP. I want to leave the \_hap piece on each of the haplotype loci so that it's clear which are SNPs and which are mhaps.

```{r}
snp <- snp %>% 
  rename_with(~ gsub(".", "_",.x, fixed = TRUE))
# make note of the current order of loci in the data set
nc <- ncol(snp)
loci <- str_replace( # expects string, pattern, replacement
  names(snp)# returns the names of the columns
  [seq(2, nc, by = 2)], # creates a sequence of even numbers starting at 2 and ending at nc, stepping by 2
  "\\.\\.\\.[0-9]+$", "") # this cleans up column names. So I don't think this will impact my dataframe at all

#  reset the locus names
names(snp)[seq(2, nc, by = 2)] <- str_c(loci, "1", sep = ".") #put a .1 at the end of the first allele
names(snp)[seq(3, nc, by = 2)] <- str_c(loci, "2", sep = ".") # put a .2 at the end of the second allele
names(snp)

# rename Sample_ID column to match mhaps. so that's indiv.ID
snp <- snp %>% 
  rename(indiv.ID = Sample_ID)
```
