---
title: "Multi to biallelic"
format: html
editor: visual
---

## Prepping SNP/mhap data for GRM

There are three key things that need to happen for the SNP and microhaplotype data to be prepped for constructing the GRM. The key thing to remember is that both data types need to be in 0, 1, and 2 allele dosage format, where dosage counts the number alleles of each type.

1.  Convert SNP data to allele dosage. I will make a function to do this, which can be reused.
2.  Count unique microhaplotypes for each mhap
3.  Convert mhaps into allele dosage, where each unique mhap is its own column.

Once both data sets are in that format, they are combined into one data set.

## Structured Plan

1.  import SNP and mhap csv's, separately.
2.  Format column names properly. I believe they should end in \_#
3.  Convert SNPs into dosages
4.  Find each unique mhap at each locus
5.  Expand into dosage columns

Since there may be a lot of columns per mhap locus, it might be beneficial to weight the alleles.

```{r}
library(here)
library(tidyverse)
```

### 1. Import and clean .csv files

```{r}
# SNP
snp_raw <- read.csv(here("Genomics/Data/PaigeAdapCapFishSNPGenotypes.raw.csv"))

# Neutral mhap
neutral_mhap <- read.csv(here("Genomics/Data/Neutral_genotypes_mhapcode_wide.csv"))

# Adaptive mhap
adap_mhap <- read.csv(here("Genomics/Data/Adaptive_haplotypes_2col_wide.csv"))
# this is the wrong file- we need the hap_code file, not haplotypes. 


```

### 2. Format/Prepare Data Frames
#### 2a. Mhap data frames
First, the mhap columns need to be coded properly in the adap_mhap file. They need to be hapcodes not haplotypes.

To do thi, I need R to count the number of unique values for each pair of haplotypes. This should be relatively straightforward- we want R to recognize two columns with the same name (excluding haplotype.1 and haplotype.2) as the same locus, then assign a number for each allele. 0's must remain 0's, as these represent loci that didn't amplify.

```{r}
# Chat GPT mostly wrote this

# this creates another version of the original data frame to play with
adap_code <- adap_mhap

# Find the base name of each locus for each column allele (remove .1/.2 suffix)
bases <- unique(sub("\\.[12]$", "", names(adap_mhap))) # sub replacestext in character string: sub(pattern, replacement, text)
# so for each unique base code (or locus), counts the number of unique values

for (b in bases) { # loop over each base

  cols <- paste0(b, c(".1", ".2")) # this recreates the column names

  # Make sure both columns (.1 and .2 exist)
  if (!all(cols %in% names(adap_mhap))) next # %in% checks whether values are present
  # all reutrns true only if every value is true
  # ! means not, so "false"
  # so then  this says, if the answer is false or missing, skip the pair

  # collect all values in a pair (excluding 0)
  vals <- setdiff(unique(unlist(adap_mhap[cols])), "0") # unlist turns cols into one long vector and unique removes duplicates - so vals contains every unique haplotype in that pair

  # Create lookup table
  lookup <- setNames(seq_along(vals), vals) # seq_along makes numbers from 1 to length of vals and setNames adds names to those numbers

  # Replace values with numbers, but keep 0 as 0
adap_code[cols] <- lapply(adap_mhap[cols], function(x) {
  out <- lookup[x]
  out[x == "0"] <- 0
  out
})
}

adap_code
#make sure 0's are maintained
sum(adap_code == 0, na.rm = TRUE)

length(unique(adap_mhap$Omy_Ch25_Six6_61287415_haplotype.2))
length(unique(adap_code$Omy_Ch25_Six6_61287415_haplotype.2))
```

Save this as a new csv:
```{r}
write.csv(adap_code, "Data/Adaptive_genotypes_mhapcode_wide.csv", row.names = FALSE)
```


Next, I need to format the column names to match the neutral mhap data frame.
I do this by renaming the neutral genotype endings so it ends in .1 and .2 to match the adaptive haplotype panel structure and snp structure.

```{r}
names(neutral_mhap) <- ifelse(
  grepl("hap1_code$", names(neutral_mhap)),
  sub("hap1_code$","haplotype.1", names(neutral_mhap)),
ifelse(
  grepl("hap2_code$", names(neutral_mhap)),
  sub("hap2_code$","haplotype.2",names(neutral_mhap)),
  names(neutral_mhap)
)
)

```

Combine data frames
```{r}
all_mhap <- merge(neutral_mhap, adap_code, by = "indiv.ID")
```


#### 2b. SNP data frames
Get rid of the sex ID column in the SNP data frame.

```{r}
snp <- snp_raw %>% 
  dplyr::select(-c(SexID,SexID.1))
```

Reformatting columns for SNPs:

```{r}
snp <- snp %>% 
  rename_with(~ gsub(".", "_",.x, fixed = TRUE))
# make note of the current order of loci in the data set
nc <- ncol(snp)
loci <- str_replace( # expects string, pattern, replacement
  names(snp)# returns the names of the columns
  [seq(2, nc, by = 2)], # creates a sequence of even numbers starting at 2 and ending at nc, stepping by 2
  "\\.\\.\\.[0-9]+$", "") # this cleans up column names. So I don't think this will impact my dataframe at all

#  reset the locus names
names(snp)[seq(2, nc, by = 2)] <- str_c(loci, "1", sep = ".") #put a .1 at the end of the first allele
names(snp)[seq(3, nc, by = 2)] <- str_c(loci, "2", sep = ".") # put a .2 at the end of the second allele
names(snp)

# rename Sample_ID column to match mhaps. so that's indiv.ID
snp <- snp %>% 
  rename(indiv.ID = Sample_ID)
```


### 3a. Combine mhaps into genotypes

We want to combine the alleles into one genotype, structured x/y.

```{r}
# this creates another version of the original data frame to play with. I'm starting with just mhaps
# Start with a new data frame based on all_mhap
mhap_geno <- data.frame(indiv.ID = all_mhap$indiv.ID) # make sure to keep ID col
# Find the unique base names, e.g. A from A.1 and A.2
bases <- unique(sub("\\.[12]$", "", names(all_mhap)))

for (b in bases) {
  
  # Get the paired column names
  cols <- paste0(b, c(".1", ".2"))
  
  # Make sure both columns exist
  if (!all(cols %in% names(all_mhap))) next
  
  # Combine the two columns, keeping 0
  mhap_geno[[b]] <- paste(
    all_mhap[[cols[1]]],
    all_mhap[[cols[2]]],
    sep = "/"
  )
}
```

The above code appears to have worked, so now I will proceed with allele dosage columns. The columns and values will be structured as follows:

###4. Convert into allele dosages 
- a unique column for each allele for each mhap. For example, if there are 7 unique mhaps at one site, then there will be 7 columns
- Each column will be formated to end with a _a, _b, etc. 
- The values will count the number of each alleles each individual has- for example:
 - 1/1 will result in a value of 2 in column a
 - 1/2 will result in a value of 1 in column a and 1 in column b
 - 2/2 will result in a value of 2 in column b and 0 in a
The key thing here is each mhap site should always add up to 2.

For sites that didn't amplify and thus have a 0 value, I replace that value with an NA. 

####4a. Mhaps -> dosages

```{r}
# modified from ChatGPT
make_haplotype_dosage <- function(x) {
  
  # ----------------------------------------------------------
  # STEP 1: Convert to character
  # ----------------------------------------------------------
  
  x <- as.character(x)
  
  
  # ----------------------------------------------------------
  # STEP 2: Split each genotype into its two haplotypes
  #
  # Example:
  #
  #     "1/2" -> "1", "2"
  #     "2/2" -> "2", "2"
  #     "0/0" -> "0", "0"
  # ----------------------------------------------------------
  
  split_haps <- strsplit(x, split = "/")
  
  
  # ----------------------------------------------------------
  # STEP 3: Identify the real haplotypes
  #
  # unlist() puts all haplotypes into one vector.
  #
  # We explicitly REMOVE "0" here because 0 means
  # no amplification, NOT a real haplotype.
  # ----------------------------------------------------------
  
  all_haps <- unlist(split_haps)
  
  haplotypes <- sort(
    unique(
      all_haps[
        !is.na(all_haps) &
        all_haps != "0"
      ]
    )
  )
  
  
  # ----------------------------------------------------------
  # STEP 4: Create an empty matrix
  #
  # One row for every individual.
  #
  # One column for every real haplotype found at this site.
  #
  # We initially fill it with 0 
  # ----------------------------------------------------------
  
  out <- matrix(
    0,
    nrow = length(x),
    ncol = length(haplotypes)
  )
  
  
  # ----------------------------------------------------------
  # STEP 5: Give the columns temporary names
  # ----------------------------------------------------------
  
  colnames(out) <- haplotypes
  
  
  # ----------------------------------------------------------
  # STEP 6: Calculate haplotype dosage
  # ----------------------------------------------------------
  
  for (j in seq_along(haplotypes)) {
    
    # Current haplotype
    h <- haplotypes[j]
    
    
    # Go through each individual
    out[, j] <- sapply(split_haps, function(z) {
      
      
      # ------------------------------------------------------
      # If the genotype is missing entirely, return NA
      # ------------------------------------------------------
      
      if (length(z) != 2 || any(is.na(z))) {
        return(NA)
      }
      
      
      # ------------------------------------------------------
      # If either haplotype is 0, the site did not amplify
      # properly.
      #
      # Rather than interpreting 0 as a real allele, mark
      # the entire site as missing.
      # ------------------------------------------------------
      
      if (any(z == "0")) {
        return(NA)
      }
      
      
      # ------------------------------------------------------
      # Count copies of the current haplotype.
      #
      # 1/1 -> 2
      # 1/2 -> 1
      # 2/2 -> 0
      # ------------------------------------------------------
      
      sum(z == h)
    })
  }
  
  
  # ----------------------------------------------------------
  # STEP 7: Rename the columns a, b, c, etc.
  # ----------------------------------------------------------
  
  colnames(out) <- letters[seq_len(ncol(out))]
  
  
  # ----------------------------------------------------------
  # STEP 8: Return the result
  # ----------------------------------------------------------
  
  return(out)
}
```

```{r}
# ------------------------------------------------------------
# APPLY TO EVERY MHAP COLUMN
# ------------------------------------------------------------
# Get the names of all columns EXCEPT indiv.ID
mhap_columns <- setdiff(
  names(mhap_geno),
  "indiv.ID"
)


# Process each mhap column
result <- lapply(mhap_columns, function(site_name) {
  
  # Convert this mhap column into haplotype dosage columns
  x <- make_haplotype_dosage(mhap_geno[[site_name]])
  
  # Add the original site name to the new column names
  #
  # For example:
  #     a   b   c
  #
  # becomes:
  #     site1_a   site1_b   site1_c
  
  colnames(x) <- paste0(
    site_name,
    "_",
    colnames(x)
  )
  
  return(x)
})


# ------------------------------------------------------------
# COMBINE ALL NEW MHAP COLUMNS
# ------------------------------------------------------------

result <- as.data.frame(
  do.call(cbind, result)
)


# ------------------------------------------------------------
# ADD indiv.ID BACK TO THE FRONT
# ------------------------------------------------------------

result <- cbind(
  indiv.ID = mhap_geno$indiv.ID,
  result
)

```

```{r}
sum(is.na(result))
```

It seems that the NA values were maintained, and so I am concluding that this worked. 

####4b. SNPs -> dosages 

###5. Combine mhaps and snps into one data frame 

The next steps, to be continued in a different file, will be to combine the 