---
title: "Multi to biallelic"
format: html
editor: visual
---

## Prepping SNP/mhap data for GRM

There are three key things that need to happen for the SNP and microhaplotype data to be prepped for constructing the GRM. The key thing to remember is that both data types need to be in 0, 1, and 2 allele dosage format, where dosage counts the number alleles of each type.

1.  Convert SNP data to allele dosage. I will make a function to do this, which can be reused.
2.  Count unique microhaplotypes for each mhap
3.  Convert mhaps into allele dosage, where each unique mhap is its own column.

Once both data sets are in that format, they are combined into one data set.

## Structured Plan

1.  import SNP and mhap csv's, separately.
2.  Format column names properly. I believe they should end in \_#
3.  Convert SNPs into dosages
4.  Find each unique mhap at each locus
5.  Expand into dosage columns

Since there may be a lot of columns per mhap locus, it might be beneficial to weight the alleles.

```{r}
library(here)
library(tidyverse)
```

### 1. Import and clean .csv files

```{r}
# SNP
snp_raw <- read.csv(here("Genomics/Data/PaigeAdapCapFishSNPGenotypes.raw.csv"))

# Neutral mhap
neutral_mhap <- read.csv(here("Genomics/Data/Neutral_genotypes_mhapcode_wide.csv"))

# Adaptive mhap
adap_mhap <- read.csv(here("Genomics/Data/Adaptive_haplotypes_2col_wide.csv"))
# this is the wrong file- we need the hap_code file, not haplotypes. 

# merged adaptive and neutral mhap data frames
all_mhap <- merge(neutral_mhap, adap_mhap, by = "indiv.ID")
```

Get rid of the sex ID column in the SNP data frame.

```{r}
snp <- snp_raw %>% 
  dplyr::select(-c(SexID,SexID.1))

?ckmrsim
```

### 1a. Converting haplotypes to hapcodes

The adaptive panel has

### 2. Format columns

Reformatting columns for SNPs:

```{r}
snp <- snp %>% 
  rename_with(~ gsub(".", "_",.x, fixed = TRUE))
# make note of the current order of loci in the data set
nc <- ncol(snp)
loci <- str_replace( # expects string, pattern, replacement
  names(snp)# returns the names of the columns
  [seq(2, nc, by = 2)], # creates a sequence of even numbers starting at 2 and ending at nc, stepping by 2
  "\\.\\.\\.[0-9]+$", "") # this cleans up column names. So I don't think this will impact my dataframe at all

#  reset the locus names
names(snp)[seq(2, nc, by = 2)] <- str_c(loci, "1", sep = ".") #put a .1 at the end of the first allele
names(snp)[seq(3, nc, by = 2)] <- str_c(loci, "2", sep = ".") # put a .2 at the end of the second allele
names(snp)

# rename Sample_ID column to match mhaps. so that's indiv.ID
snp <- snp %>% 
  rename(indiv.ID = Sample_ID)
```

Reformatting columns for mhaps:

I left off here. The column names need to match the format of the SNP. I want to leave the \_hap piece on each of the haplotype loci so that it's clear which are SNPs and which are mhaps.

```{r}
# make note of the current order of loci in the data set
nc2 <- ncol(all_mhap)
loci2 <- str_replace( # expects string, pattern, replacement
  names(all_mhap)# returns the names of the columns
  [seq(2, nc2, by = 2)], # creates a sequence of even numbers starting at 2 and ending at nc, stepping by 2
  "\\.\\.\\.[0-9]+$", "") # this cleans up column names. So I don't think this will impact my dataframe at all

#  reset the locus names
names(all_mhap)[seq(2, nc, by = 2)] <- str_c(loci2, "1", sep = ".") #put a .1 at the end of the first allele
names(all_mhap)[seq(3, nc, by = 2)] <- str_c(loci2, "2", sep = ".") # put a .2 at the end of the second allele
names(all_mhap)

```

### 3a. Combine mhaps into genotypes

We want to combine the alleles into one genotype, structured x/y. First, the mhap columns need to be coded properly.

I need R to count the number of unique values for each pair of haplotypes. This should be relatively straightforward- we want R to recognize two columns with the same name (excluding haplotype.1 and haplotype.2) as the same locus.

This below code works (thanks chat), but I think I need to go back and tell it to leave 0's as 0's.

```{r}
# Chat GPT mostly wrote this

# this creates another version of the original data frame to play with
adap_code <- adap_mhap

# Find the base name of each locus for each column allele (remove .1/.2 suffix)
bases <- unique(sub("\\.[12]$", "", names(adap_mhap))) # sub replacestext in character string: sub(pattern, replacement, text)
# so for each unique base code (or locus), counts the number of unique values

for (b in bases) { # loop over each base

  cols <- paste0(b, c(".1", ".2")) # this recreates the column names

  # Make sure both columns (.1 and .2 exist)
  if (!all(cols %in% names(adap_mhap))) next # %in% checks whether values are present
  # all reutrns true only if every value is true
  # ! means not, so "false"
  # so then  this says, if the answer is false or missing, skip the pair

  # collect all values in a pair (excluding 0)
  vals <- setdiff(unique(unlist(adap_mhap[cols])), "0") # unlist turns cols into one long vector and unique removes duplicates - so vals contains every unique haplotype in that pair

  # Create lookup table
  lookup <- setNames(seq_along(vals), vals) # seq_along makes numbers from 1 to length of vals and setNames adds names to those numbers

  # Replace values with numbers, but keep 0 as 0
adap_code[cols] <- lapply(adap_mhap[cols], function(x) {
  out <- lookup[x]
  out[x == "0"] <- 0
  out
})
}

adap_code
#make sure 0's are maintained
sum(adap_code == 0, na.rm = TRUE)

length(unique(adap_mhap$Omy_Ch25_Six6_61287415_haplotype.2))
length(unique(adap_code$Omy_Ch25_Six6_61287415_haplotype.2))
```

```{r}
write.csv(adap_code, "Data/Adaptive_genotypes_mhapcode_wide.csv", row.names = FALSE)
```

### 3b. Combine SNPs into genotypes

Ok so I didn't make a mistake- the file I loaded in was the haplotypes, but I need the hap_code file and I don't think this has been made yet.
